### 2017-06-16 manuscript/manuscript.Rmd
initial commit

---
title: Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
author: "Ross Cunning"
date: "6/7/2017"
output: html_document
---

Title ideas:

Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals

### 2017-06-16 supp/supp.Rmd
initial commit

---
title: "Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals"
output: html_document
---

```{r setup, include=FALSE}
# Load libraries
library(MASS)
library(effects)
library(reshape2)
library(scales)
library(lme4)
library(lmerTest)
library(lsmeans)
library(lattice)
library(dplyr)

# Set knit options
knitr::opts_chunk$set(warning=F, message=F, fig.height=5, fig.width=5)
knitr::opts_knit$set(root.dir = "..")

# Polygon drawing function
addpoly <- function(x, y1, y2, col=alpha("lightgrey", 0.8), ...){
  ii <- order(x)
  y1 <- y1[ii]
  y2 <- y2[ii]
  x <- x[ii]
  polygon(c(x, rev(x)), c(y1, rev(y2)), col=col, border=NA, ...)
}
```

# Import data
```{r load_data}
alldata <- read.csv("data/master.csv")
alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
```

# Analyze Fv/Fm using the repeat bleaching data for *M. cavernosa*
```{r}

# Get data frame with all Ofav and Ssid data
SsOf <- subset(alldata, Species!="Mc")

# Get data frame with Mcav repeat bleaching data (=second bleaching)
Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
## Re-level Mcav data for merging with Ofav and Ssid data
## Set data after recovery from first bleaching to the "initial" factor levels
Mc$PropD.init <- Mc$PropD.recov  
Mc$MaxY.init <- Mc$MaxY.recov
## Set data after the second bleaching to the "bleached" factor levels
Mc$BleachLvl <- as.character(Mc$BleachLvl2)
Mc$PropD.bleach <- Mc$PropD.bleach2
Mc$MaxY.bleach <- Mc$MaxY.bleach2
## Set BleachLvl to "Med" to match Ofav and Ssid factor level
Mc$BleachLvl[Mc$BleachLvl=="Heat"] <- "Med"
## Set data after second recovery to the "recove#d95f02" factor levels
Mc$PropD.recov <- Mc$PropD.recov2
Mc$MaxY.recov <- Mc$MaxY.recov2

# Merge data from all three coral species together
bleach3c <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "Total.init", "PropD.init", "MaxY.init", "BleachLvl", "Total.bleach", "PropD.bleach", "MaxY.bleach")]
bleach3c <- droplevels(bleach3c)
```

### Initial Fv/Fm before stress
```{r MaxY.init.3c, fig.width=7, fig.height=4}

# Plot Fv/Fm vs. PropD by Species
xyplot(MaxY.init ~ PropD.init | Species, bleach3c, 
       main="Initial Fv/Fm by species",
       xlab="Initial proportion clade D", ylab="Initial Fv/Fm")

# Fit mixed model with fixed effects of Species and proportion clade D
maxyinit.m <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
                   data=bleach3c) 
anova(maxyinit.m)  # Assess significance of Species

# Calculate mean Fv/Fm for each species at PropD=0 w/ pairwise comparisons
maxyinit.lsm <- lsmeans(maxyinit.m, "Species", 
                        at=list(PropD.init=0))  
cld(maxyinit.lsm) # View lsmean comparisons
plot(maxyinit.lsm, main="Initial Fv/Fm with no clade D", horizontal=F)

# Effect of proportion clade D on initial Fv/Fm
maxyinit.m.lst <- lstrends(maxyinit.m, specs="Species", var="PropD.init")
cld(maxyinit.m.lst) # Slopes of fit for each species
plot(Effect(c("PropD.init", "Species"), maxyinit.m))  # Plot model fit

# Difference in Fv/Fm with 100% clade D?
maxyinit.lsm <- lsmeans(maxyinit.m, "Species", at=list(PropD.init=1))
cld(maxyinit.lsm)
plot(maxyinit.lsm, main="Initial Fv/Fm with 100% clade D", horizontal=F)
```

* Higher amounts of clade D #d95f02uce Fv/Fm, with this effect being stronger in *O. faveolata* compa#d95f02 to *S. siderea* and *M. cavernosa*. Fv/Fm with 100% D is lower in *M. cavernosa* than the other two species.

```{r MaxY.bleach.3c, fig.width=7, fig.height=4}

# Fit mixed model with fixed effects of Species and proportion clade D and duration
maxybleach.m <- lmer(MaxY.bleach ~ PropD.bleach * Species * BleachLvl + (1|Colony),
                     data=bleach3c)
select <- lmerTest::step(maxybleach.m)$model
maxybleach.m <- lmer(formula(select), data=bleach3c)
anova(maxybleach.m)

# Calculate mean Fv/Fm for each species at PropD=0 and BleachLvl=Med
maxybleach.0.lsm <- lsmeans(maxybleach.m, "Species", 
                          at=list(PropD.bleach=0, BleachLvl="Med"))  
cld(maxybleach.0.lsm) # View lsmean comparisons
plot(maxybleach.0.lsm, main="Bleached Fv/Fm with 0% clade D", horizontal=F)
maxybleach.0.summ <- summary(maxybleach.0.lsm)

# look at slopes and 100% clade D Fv/Fm
maxybleach.1.lsm <- lsmeans(maxybleach.m, "Species", 
                          at=list(PropD.bleach=1, BleachLvl="Med"))  
cld(maxybleach.1.lsm) # View lsmean comparisons
plot(maxybleach.1.lsm, main="Bleached Fv/Fm with 100% clade D", horizontal=F)
maxybleach.1.summ <- summary(maxybleach.1.lsm)

# Test differences in slopes
maxybleach.lst <- lstrends(maxybleach.m, specs="Species", var="PropD.bleach")
cld(maxybleach.lst) # Slopes of fit for each species
#plot(Effect(c("PropD.bleach", "Species"), maxybleach.m))  # Plot model fit
```

* With no clade D, *O. faveolata* maintained significantly higher Fv/Fm (`r maxybleach.0.summ$lsmean[1]`±`r maxybleach.0.summ$SE[1]`) than *S. siderea*, while *M. cavernosa* was also lower but not significantly different.
* *O. faveolata* and *S. siderea* with 100% clade D had the same Fv/Fm (0.31) under heat stress, while *M. cavernosa* with clade D had a significantly higher Fv/Fm (0.38). 
* As the proportion of clade D went down, Fv/Fm went down > 3 times faster in *S. siderea* and almost 5 times faster in *M. cavernosa* compa#d95f02 to *O. faveolata*, suggesting their other symbionts differ in thermal sensitivity.

```{r, eval=T}
# Melt data frame so that "time" (i.e., initial vs. bleached) can be used as factor
bleachmelt <- melt(bleach3c, id.vars = c("Species", "Colony", "Core", "BleachLvl"))
bleachmelt <- cbind(bleachmelt, 
                    colsplit(bleachmelt$variable, pattern="\\.", 
                             names=c("resp", "time")))
bleachdat <- dcast(bleachmelt, 
                   formula = Species + Colony + Core + BleachLvl + time ~ resp, value.var="value")
bleachdat$time <- factor(bleachdat$time)

# Test of slopes of FvFm ~ PropD are different for different BleachLvls
testmod <- lmer(MaxY ~ PropD * Species * BleachLvl + (1|Colony), 
                data=subset(bleachdat, time=="bleach" & Species!="Mc"))
anova(testmod)
plot(Effect(c("PropD", "BleachLvl", "Species"), testmod), multiline=T, z.var="BleachLvl")
slopes <- lstrends(testmod, specs=c("BleachLvl", "Species"), var="PropD")
cld(contrast(slopes, by="Species"))
## No difference in photochemical advantage of D under stress among different BleachLvl's. Therefore, can omit BleachLvl from statistical model for calculating photochemical advantage of D

# Fit model for Fv/Fm
bleachmod <- lmer(MaxY ~ PropD * time * Species + (1|Colony), bleachdat)
anova(bleachmod)
select <- lmerTest::step(bleachmod)$model
bleachmod <- lmer(formula(select), data=bleachdat)
anova(bleachmod)

plot(Effect(c("PropD", "time", "Species"), bleachmod), multiline=T, ci.style="bands")

# Compare slopes of Fv/Fm vs. PropD among species by time
lst <- lstrends(bleachmod, specs=c("Species", "time"), var="PropD")
cld(contrast(lst, by="time"))
plot(lst, horizontal=F)

# Make data frame to collect fitted values and confidence intervals
pred <- expand.grid(Species=c("Of", "Ss", "Mc"), 
                    PropD=seq(0,1,len=101), 
                    time=c("init", "bleach"))
# Get fitted Fv/Fm values from model
pred$fit <- predict(bleachmod, pred, re.form=NA)

# Bootstrap model to get confidence intervals on fitted values
bootfit <- bootMer(bleachmod, FUN=function(x) predict(x, pred, re.form=NA), nsim=999)
# Extract 90% confidence interval on predicted values
pred$lci <- apply(bootfit$t, 2, quantile, 0.05)
pred$uci <- apply(bootfit$t, 2, quantile, 0.95)
```

```{r Fig1}
png("figures/Fig1.png", width=4, height=2.5, units="in", res=300)
par(mfrow=c(1, 3), mar=c(3,3,1.1,0), mgp=c(1.5, 0.1, 0), 
    tcl=0.2, cex.axis=0.8, xaxs="i")
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="Fv/Fm")
mtext("A.", adj=-0.2, font=2)
with(subset(pred, Species=="Of" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#1b9e77", 0.5))
})
with(subset(pred, Species=="Of" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#1b9e77", 0.5))
})

par(mar=c(3,1.5,1.1,1.5))
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="")
mtext("B.", adj=-0.2, font=2)
with(subset(pred, Species=="Ss" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#7570b3", 0.5))
})
with(subset(pred, Species=="Ss" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#7570b3", 0.5))
})

par(mar=c(3,0,1.1,3))
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="")
mtext("C.", adj=-0.2, font=2)
with(subset(pred, Species=="Mc" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#d95f02", 0.5))
})
with(subset(pred, Species=="Mc" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#d95f02", 0.5))
})
dev.off()
```
![Figure 1.](../figures/Fig1.png)

```{r Fig2}
# Plot Advantage of D vs. other symbiont in each coral under
# normal conditions and under heat stress (in Fv/Fm units)

# Get ADH and ADA values (Advantage of D under Heat and under Ambient conditions)
AD <- summary(lst)
ADm <- melt(AD)
ADm$time <- factor(ADm$time, levels=c("init", "bleach"))
ADf <- as.matrix(dcast(subset(ADm, variable=="PropD.trend"), time ~ Species)[,-1])
ADu <- as.matrix(dcast(subset(ADm, variable=="upper.CL"), time ~ Species)[, -1])
ADl <- as.matrix(dcast(subset(ADm, variable=="lower.CL"), time ~ Species)[, -1])

png("figures/Fig2.png", width=3, height=3, units="in", res=300)
par(mar=c(1.8,3.2,1,0.8), mgp=c(1.5, 0.25, 0), tcl=-0.2, cex.axis=0.8)
bars <- barplot(ADf, beside = T, ylim=c(-0.2,0.2),
                names.arg=c("Ofav", "Ssid", "Mcav"),
                col=c("#1b9e77", alpha("#1b9e77", 0.3),
                      "#7570b3", alpha("#7570b3", 0.3),
                      "#d95f02", alpha("#d95f02", 0.3)))
arrows(matrix(bars), matrix(ADl), matrix(bars), matrix(ADu), 
       length = 0.05, lwd=1, xpd=NA, code=3, angle=90)
mtext(side=2, "Photochemical advantage of clade D", 
      line=1.5, cex=0.8)
#legend("topleft", pch=22, pt.bg=c("gray80", "gray20"), y.intersp=1.2,
#       pt.cex=2, cex=0.6, inset=c(0.02, -0.05), xpd=NA, bty = "n", 
#       legend = c(expression("Heat stress"~"(A"["Heat"]^"D"~")"),
#                  expression("Ambient"~"(A"["Amb"]^"D"~")")))
abline(h=0)
dev.off()
```
![Figure 2](../figures/Fig2.png)

# Symbiont shuffling after bleaching
```{r}
recovdat <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "Total.init", "PropD.init", "MaxY.init", "BleachLvl", "RecovTemp", "Total.recov", "PropD.recov", "MaxY.recov")]

propDmod <- glm(PropD.recov ~ PropD.init + Species*BleachLvl*RecovTemp,
                 data=recovdat, family=quasibinomial)
summary(propDmod)  # Model summary
dropterm(propDmod, test="F")  # Partial F-tests: Results in Table 1

# Get fitted values for PropD.recov, averaging across RecovTemp levels
eff <- effect(c("PropD.init", "Species", "BleachLvl"), propDmod, 
              xlevels=list(PropD.init=seq(0, 1, 0.01)))

# Get all fitted values and subsets for each bleaching level
res <- droplevels(data.frame(eff))
res$Species <- factor(res$Species, levels=c("Of", "Ss", "Mc"))
res.Med <- subset(data.frame(eff), BleachLvl=="Med")
res.Low <- subset(data.frame(eff), BleachLvl=="Low")
res.High <- subset(data.frame(eff), BleachLvl=="High")

# Get AUC for fitted values, lower and upper confidence limits
auc <- aggregate(res[, c("fit", "lower", "upper")], 
                 by=list(Species=res$Species, BleachLvl=res$BleachLvl), 
                 FUN=function(x) (mean(x)-0.5)/0.5)
auc.list <- split(auc, list(auc$BleachLvl))
```

```{r Fig3}
# Figure 3
png("figures/Fig3.png", width=7, height=3.5, units="in", res=300)
par(mfrow=c(1, 2), mar=c(3,3,1,1), mgp=c(1.5, 0.1, 0), tcl=0.2, cex.axis=0.8)
## Figure 3A: Final vs. initial proportion clade D
plot(NA, xlim=c(0,1), ylim=c(0,1), xaxs="i", yaxs="i",
     xlab="Prop. D before bleaching",
     ylab="Prop. D after recovery")
# Plot lines for low bleaching
with(subset(res.Low, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Low, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Low, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
abline(0,1, lty=1, lwd=0.5)
# Plot lines for Med bleaching
with(subset(res.Med, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Med, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Med, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
# Plot lines for High bleaching
with(subset(res.High, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.High, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.High, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
legend("bottomright", bty="n", lty=c(1,2,3), lwd=2, inset=c(0.05, 0),
       legend=c("7 days", "10 days", "14 days"), cex=0.7)
legend("bottomright", bty="n", pch=15, inset=c(0.04, 0.175),
       legend=c("Mcav", "Ssid", "Ofav"), cex=0.7, pt.cex=1.5, x.intersp=1.5,
       col=c("#d95f02", "#7570b3", "#1b9e77"))
#text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "A.", font=2, xpd=NA)
#text(-0.2, par("usr")[4], "A.", font=2, xpd=NA)
mtext("A.", adj=-0.2, font=2)

# Figure 3B: AUC for each species at each BleachLvl
plot(NA, ylim=c(-1,1), xlim=c(0.5,3.5), xaxs="i", yaxs="i", xaxt="n", 
     xlab="Species", ylab="Symbiont shuffling")
axis(side=1, at=1:3, labels=c("Of", "Ss", "Mc"))
abline(h=0, lty=1, lwd=0.5)
with(auc.list$Low, {
  arrows(0.8:2.8, lower, 0.8:2.8, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(0.8:2.8, fit, pch=15, col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
with(auc.list$Med, {
  arrows(1:3, lower, 1:3, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(1:3, fit, pch=17, col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
with(auc.list$High, {
  arrows(1.2:3.2, lower, 1.2:3.2, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(1.2:3.2, fit, pch=16,  col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       pch=c(15, 17, 16), col="black", cex=0.7,
       legend=c("7 days", "10 days", "14 days"))

arrows(par("usr")[2]*0.95, c(0.02, -0.02), 
       par("usr")[2]*0.95, c(0.2, -0.2), length=0.1)
text(par("usr")[2]*0.95, c(0.08, -0.1), pos=2, cex=0.7,
     labels=c("more clade D", "less clade D"))
#text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "B.", font=2, xpd=NA)
mtext("B.", adj=-0.2, font=2)
dev.off()
```
![Figure 3.](../figures/Fig3.png)

```{r Fig4}

# Model shuffling as a function of photochemical advantage and heat stress

# Get AD (photochemical Advantage of D, =slopes of Fv/Fm~propD lines)
AD <- summary(lst)[,1:3]
df <- merge(AD, auc)
## ADA = Advantage of D under Ambient conditions (not relative -- in Fv/Fm units)
df$ADA <- with(subset(AD, time=="init"), 
                      PropD.trend[match(df$Species, Species)])
df$ADH <- with(subset(AD, time=="bleach"), 
                      PropD.trend[match(df$Species, Species)])
# Change column name for fitted shuffling response to 'shuffle'
colnames(df) <- stringr::str_replace(colnames(df), "fit", "shuffle")
# Convert BleachLvl factor to # days of heat stress
df$days <- ifelse(df$BleachLvl=="Low", 7, ifelse(df$BleachLvl=="Med", 10, 14))
# Rescale shuffling response range from [-1,1] to [0,1]
df$shuffle.scaled <- scales::rescale(df$shuffle, from=c(-1, 1), to=c(0, 1))
# Create variables for products of Fv/Fm differences (AD metrics) and days of heat
df$ADH.days <- df$ADH * df$days
df$ADA.days <- df$ADA * df$days

# P#d95f02ict shuffling response based on relative Fv/Fm value and BleachLvl
## relative Fv/Fm at ambient temperatures ("normal") is perfectly correlated with relative Fv/Fm at stress temperatures ("stress") so only one variable can be used in model.  Or, should the slope of the Fv/Fm line be used, rather than the "relative difference"? (makes very little difference whether relative or slope is used.) *slope needs to be used because relative is different for diff BleachLvl levels, but slope is the same!

png("figures/Fig4.png", width=7, height=3.5, units="in", res=300)
par(mfrow=c(1,2), mgp=c(1.5,0.1,0), mar=c(3,3,1,1), tcl=0.2, cex.axis=0.8)
# Orthogonal model
glmod <- glm(shuffle.scaled ~ ADH + days, data=df, family="quasibinomial")
anova(glmod, test="F")

# Get fitted values and plot shuffling ~ days for different relative advantages of D
stresslevs <- c(0.033, 0.05, 0.07, 0.103, 0.163)
cols <- c("#1b9e77", "gray40", "gray40", "#7570b3", "#d95f02")
resdf <- expand.grid(days=7:14, ADH=stresslevs)
resdf$fit <- predict(glmod, newdata=resdf, type="response")
resdf$fit.bt <- rescale(resdf$fit, from=c(0,1), to=c(-1,1))
plot(NA, xlim=c(7,14), ylim=c(-1,1), yaxs="r",
     xlab="Days of heat stress",
     ylab="Symbiont shuffling")
mtext("A.", adj=-0.2, font=2)
reslist <- split(resdf, f=resdf$ADH)
for (i in 1:length(reslist)) {
  with(reslist[[i]], lines(fit.bt ~ days, col=cols[i], lwd=2, xpd=NA))
}
abline(h=0, lwd=0.5)
# Get positions for text annotations
angles <- unlist(lapply(reslist, function(x) with(x, fit.bt[2]-fit.bt[1])))
heights <- unlist(lapply(reslist, function(x) with(x, fit.bt[1])))
ADHeat <- expression("A"["Heat"]^"D"~"=")
for (i in 1:length(stresslevs)) {
  text(labels=bquote(.(parse(text=paste(ADHeat,"~",stresslevs[i],sep=""))) ),
       x=7, y=heights[i]-0.06, adj=0, srt=angles[i]*200, cex=0.6)
}
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       lty=1, lwd=2, col=c("#d95f02", "#7570b3", "#1b9e77"), cex=0.7,
       legend=c("Mcav", "Ssid", "Ofav"))

## Panel B
glmod2 <- glm(shuffle.scaled ~ ADH.days, family="quasibinomial", data=df)
predicted <- predict(glmod2, 
                     newdata=data.frame(ADH.days=seq(0,3,0.1)), 
                     type="response")
plot(df$shuffle ~ df$ADH.days, ylim=c(-1,1), yaxs="r", xpd=NA,
     pch=c(16,15,17)[as.numeric(as.factor(df$BleachLvl))],
     xlab=expression("A"["Heat"]^"D"~"x (Days of heat stress)"),
     ylab="", mgp=c(1.8,0.1,0),
     col=c("#d95f02", "#1b9e77", "#7570b3")[as.numeric(factor(df$Species))])
mtext("Symbiont shuffling", side=2, line=1.5)
mtext("B.", adj=-0.2, font=2)
lines(seq(0,3,0.1), rescale(predicted, from=c(0,1), to=c(-1,1)), lwd=1)
abline(h=0, lwd=0.5)
text(par("usr")[2]*0.95, c(0.08, -0.1), pos=2, cex=0.7,
     labels=c("more clade D", "less clade D"))
arrows(par("usr")[2]*0.95, c(0.02, -0.02), 
       par("usr")[2]*0.95, c(0.2, -0.2), length=0.1)
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       pch=c(15, 17, 16), col="black", cex=0.7,
       legend=c("7 days", "10 days", "14 days"))
dev.off()
```
![Figure 4.](../figures/Fig4.png)

# Analyze data using initial *M. cavernosa* bleaching experiment
# Analyze photochemical efficiency (Fv/Fm)

### Initial Fv/Fm before stress
```{r MaxY.init, fig.width=7, fig.height=4}

# Plot Fv/Fm vs. PropD by Species
xyplot(MaxY.init ~ PropD.init | Species, alldata, 
       main="Initial Fv/Fm by species",
       xlab="Initial proportion clade D", ylab="Initial Fv/Fm")

# Fit mixed model with fixed effects of Species and proportion clade D
maxyinit.m <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
                   data=alldata) 
anova(maxyinit.m)  # Assess significance of Species

# Calculate mean Fv/Fm for each species at PropD=0 w/ pairwise comparisons
maxyinit.lsm <- lsmeans(maxyinit.m, "Species", at=list(PropD.init=0))  
cld(maxyinit.lsm) # View lsmean comparisons
plot(maxyinit.lsm, main="Initial Fv/Fm with no clade D", horizontal=F)

# Effect of proportion clade D on initial Fv/Fm
## (Excludes *M. cavernosa*, which did not host clade D before stress)

# Look at distribution of PropD within each colony
#histogram(~ PropD.init | Colony, data=subset(alldata, Species!="Mc"))
# Not enough variation in PropD within colonies to be able to use random slopes

# Fit mixed model with fixed effect of Species and PropD, random effect of Colony
maxyinit.m2 <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
                    data=subset(alldata, Species!="Mc"))
anova(maxyinit.m2)  # Assess significance of species and propD interaction
maxyinit.m2.lst <- lstrends(maxyinit.m2, specs="Species", var="PropD.init")
cld(maxyinit.m2.lst) # Slopes of fit for each species
plot(Effect(c("PropD.init", "Species"), maxyinit.m2))  # Plot model fit

# Difference in Fv/Fm between Ofav and Ssid with 100% clade D?
maxyinit.lsm <- lsmeans(maxyinit.m2, "Species", at=list(PropD.init=1))
cld(maxyinit.lsm)
plot(maxyinit.lsm, main="Initial Fv/Fm with 100% clade D", horizontal=F)
```

* Higher amounts of clade D #d95f02uce Fv/Fm, with this effect being stronger in *O. faveolata* compa#d95f02 to *S. siderea*. However, Fv/Fm with 100% D is similar in both species, indicating that it is the higher Fv/Fm of the non-clade-D symbiont of *O. faveolata* that is driving this trend.

### Fv/Fm after heat stress
```{r, fig.width=7, fig.height=4}

# Subset corals bleached by heat
## Corals bleached by heat for 10 days only
heat10day <- subset(alldata, BleachLvl=="Med" | BleachLvl=="Heat")
## All corals bleached by heat (7, 10, and 14 days for Ofav and Ssid)
heatbleach <- subset(alldata, !BleachLvl %in% c("DCMU", "Control"))
## Set BleachLvl for heated Mcav to "Med" (=10 days of heat) to match Ofav and Ssid
heatbleach[heatbleach$Species=="Mc", "BleachLvl"] <- "Med"

# Plot bleached Fv/Fm vs. bleached proportion clade D by species
xyplot(MaxY.bleach ~ PropD.bleach | Species, data=heatbleach)

# Look at distribution of PropD after bleaching within each colony
# histogram(~ PropD.bleach | Colony, data=heatbleach)
## Many colonies do not have a good spread of proportion clade D. Therefore, allowing random slopes for each colony when modeling responses as a function of proportion clade D will give very high variance (i.e., there is not enough data to constrain slopes within colonies). Therefore, we will model responses with random intercepts only.

# Fit mixed model with fixed effects of Species and proportion clade D and duration
maxybleach.m <- lmer(MaxY.bleach ~ PropD.bleach * Species * BleachLvl + (1|Colony),
                     data=heatbleach)
select <- lmerTest::step(maxybleach.m)$model
maxybleach.m <- lmer(formula(select), data=heatbleach)
anova(maxybleach.m)
## Not there is NO significant interaction between PropD and BleachLvl, indicating that the duration of heat stress did not influence the relationship between PropD and Fv/Fm. Therefore, we can use data from all three BleachLvl's (for Ofav and Ssid) to compare the effect of PropD on Fv/Fm among species.

# Calculate mean Fv/Fm for each species at PropD=0 and BleachLvl=Med
maxybleach.0.lsm <- lsmeans(maxybleach.m, "Species", 
                          at=list(PropD.bleach=0, BleachLvl="Med"))  
cld(maxybleach.0.lsm) # View lsmean comparisons
plot(maxybleach.0.lsm, main="Bleached Fv/Fm with 0% clade D", horizontal=F)
maxybleach.0.summ <- summary(maxybleach.0.lsm)

# Re-fit model excluding Mcav to look at slopes and 100% clade D Fv/Fm
maxybleach.m.noMc <- update(maxybleach.m, data=subset(heatbleach, Species!="Mc"))

maxybleach.1.lsm <- lsmeans(maxybleach.m.noMc, "Species", 
                          at=list(PropD.bleach=1, BleachLvl="Med"))  
cld(maxybleach.1.lsm) # View lsmean comparisons
plot(maxybleach.1.lsm, main="Bleached Fv/Fm with 100% clade D", horizontal=F)
maxybleach.1.summ <- summary(maxybleach.1.lsm)

# Test differences in slopes
maxybleach.lst <- lstrends(maxybleach.m.noMc, specs="Species", var="PropD.bleach")
cld(maxybleach.lst) # Slopes of fit for each species
plot(Effect(c("PropD.bleach", "Species"), maxybleach.m.noMc))  # Plot model fit
```

* With no clade D, *O. faveolata* maintained significantly higher Fv/Fm (`r maxybleach.0.summ$lsmean[1]`±`r maxybleach.0.summ$SE[1]`) than the other two species under heat stress.
* *O. faveolata* and *S. siderea* with 100% clade D had the same Fv/Fm (0.31) under heat stress, suggesting that clade D performs the same in both host species under stress. As the proportion of clade D went down, Fv/Fm went down > 3 times faster in *S. siderea* compa#d95f02 to *O. faveolata*, suggesting their other symbionts differ in thermal sensitivity (*M. cavernosa* still did not have clade D at this point).

### 2017-06-16 R/Fig1.R

@@ -1,0 +1,49 @@
+library(lme4)
+library(reshape2)
+library(effects)
+library(lsmeans)
+
+# Load data -----
+rm(list=ls())
+alldata <- read.csv("data/master.csv")
+alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
+alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
+
+# Init Fv/Fm for Mcav during second heat bleaching (=same as recov from 1st)
+Mc <- subset(alldata, Species=="Mc" & BleachLvl2=="Heat")
+#plot(MaxY.recov ~ PropD.recov, Mc)
+mcmaxyrecov.m <- lmer(MaxY.recov ~ PropD.recov + (PropD.recov|Colony), Mc)
+anova(mcmaxyrecov.m)
+#plot(Effect("PropD.recov", mcmaxyrecov.m))
+#plot(MaxY.bleach2 ~ PropD.bleach2, Mc)
+mcmaxybleach2.m <- lmer(MaxY.bleach2 ~ PropD.bleach2 + (PropD.recov|Colony), Mc)
+#plot(Effect("PropD.bleach2", mcmaxybleach2.m))
+
+# How much is Fv/Fm reduced in Mcav because they were previously bleached? - can't tell because changed to D - confounded
+
+# Get data frame with Of, Ss 10-day bleach  (Med bleach only or all bleachlvls pooled?)
+#SsOf <- subset(alldata, BleachLvl=="Med" & Species!="Mc")
+SsOf <- subset(alldata, Species!="Mc")
+# Get data frame with Mc 10-day bleach (from repeat bleaching experiment)
+Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
+Mc$PropD.init <- Mc$PropD.recov  # Recov from initial bleach = init for repeat bleach / for merging with Ss, Of
+Mc$MaxY.init <- Mc$MaxY.recov
+Mc$PropD.bleach <- Mc$PropD.bleach2  # repeat bleach = bleach / for merging with Ss, Of
+Mc$MaxY.bleach <- Mc$MaxY.bleach2
+# Merge three species together (from two separate but comparable bleaching experiments)
+bleach10day <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "PropD.init", "MaxY.init", "PropD.bleach", "MaxY.bleach")]
[…]

### 2017-06-16 R/Fig2.R

@@ -1,0 +1,87 @@
+
+library(lme4)
+library(reshape2)
+library(effects)
+library(lsmeans)
+
+# Load data -----
+rm(list=ls())
+alldata <- read.csv("data/master.csv")
+alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
+alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
+
+
+
+
+
+
+
+# Init Fv/Fm for Mcav during second heat bleaching (=same as recov from 1st)
+Mc <- subset(alldata, Species=="Mc" & BleachLvl2=="Heat")
+#plot(MaxY.recov ~ PropD.recov, Mc)
+mcmaxyrecov.m <- lmer(MaxY.recov ~ PropD.recov + (PropD.recov|Colony), Mc)
+anova(mcmaxyrecov.m)
+#plot(Effect("PropD.recov", mcmaxyrecov.m))
+#plot(MaxY.bleach2 ~ PropD.bleach2, Mc)
+mcmaxybleach2.m <- lmer(MaxY.bleach2 ~ PropD.bleach2 + (PropD.recov|Colony), Mc)
+#plot(Effect("PropD.bleach2", mcmaxybleach2.m))
+
+# How much is Fv/Fm reduced in Mcav because they were previously bleached? - can't tell because changed to D - confounded
+
+# Get data frame with Of, Ss 10-day bleach  (Med bleach only or all bleachlvls pooled?)
+#SsOf <- subset(alldata, BleachLvl=="Med" & Species!="Mc")
+SsOf <- subset(alldata, Species!="Mc")
+# Get data frame with Mc 10-day bleach (from repeat bleaching experiment)
+Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
+Mc$PropD.init <- Mc$PropD.recov  # Recov from initial bleach = init for repeat bleach / for merging with Ss, Of
+Mc$MaxY.init <- Mc$MaxY.recov
+Mc$PropD.bleach <- Mc$PropD.bleach2  # repeat bleach = bleach / for merging with Ss, Of
+Mc$MaxY.bleach <- Mc$MaxY.bleach2
+# Merge three species together (from two separate but comparable bleaching experiments)
+bleach10day <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "PropD.init", "MaxY.init", "PropD.bleach", "MaxY.bleach")]
[…]

### 2017-06-16 supp/supp.Rmd
add figure legends, trim extra

@@
 ```{r load_data}
+# Load all data from file
 alldata <- read.csv("data/master.csv")
@@
 alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
-```
-
-# Analyze Fv/Fm using the repeat bleaching data for *M. cavernosa*
-```{r}
 
@@
 bleach3c <- droplevels(bleach3c)
-```
 
-### Initial Fv/Fm before stress
-```{r MaxY.init.3c, fig.width=7, fig.height=4}
-
-# Plot Fv/Fm vs. PropD by Species
-xyplot(MaxY.init ~ PropD.init | Species, bleach3c, 
-       main="Initial Fv/Fm by species",
-       xlab="Initial proportion clade D", ylab="Initial Fv/Fm")
-
-# Fit mixed model with fixed effects of Species and proportion clade D
-maxyinit.m <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
-                   data=bleach3c) 
-anova(maxyinit.m)  # Assess significance of Species
-
-# Calculate mean Fv/Fm for each species at PropD=0 w/ pairwise comparisons
-maxyinit.lsm <- lsmeans(maxyinit.m, "Species", 
-                        at=list(PropD.init=0))  
-cld(maxyinit.lsm) # View lsmean comparisons
-plot(maxyinit.lsm, main="Initial Fv/Fm with no clade D", horizontal=F)
-
-# Effect of proportion clade D on initial Fv/Fm
-maxyinit.m.lst <- lstrends(maxyinit.m, specs="Species", var="PropD.init")
-cld(maxyinit.m.lst) # Slopes of fit for each species
-plot(Effect(c("PropD.init", "Species"), maxyinit.m))  # Plot model fit
-
-# Difference in Fv/Fm with 100% clade D?
-maxyinit.lsm <- lsmeans(maxyinit.m, "Species", at=list(PropD.init=1))
-cld(maxyinit.lsm)
-plot(maxyinit.lsm, main="Initial Fv/Fm with 100% clade D", horizontal=F)
-```
-
-* Higher amounts of clade D #d95f02uce Fv/Fm, with this effect being stronger in *O.
-faveolata* compa#d95f02 to *S.
-siderea* and *M.
-cavernosa*.
-Fv/Fm with 100% D is lower in *M.
-cavernosa* than the other two species.
-
-```{r MaxY.bleach.3c, fig.width=7, fig.height=4}
-
-# Fit mixed model with fixed effects of Species and proportion clade D and duration
-maxybleach.m <- lmer(MaxY.bleach ~ PropD.bleach * Species * BleachLvl + (1|Colony),
-                     data=bleach3c)
-select <- lmerTest::step(maxybleach.m)$model
-maxybleach.m <- lmer(formula(select), data=bleach3c)
-anova(maxybleach.m)
-
-# Calculate mean Fv/Fm for each species at PropD=0 and BleachLvl=Med
-maxybleach.0.lsm <- lsmeans(maxybleach.m, "Species", 
-                          at=list(PropD.bleach=0, BleachLvl="Med"))  
-cld(maxybleach.0.lsm) # View lsmean comparisons
-plot(maxybleach.0.lsm, main="Bleached Fv/Fm with 0% clade D", horizontal=F)
-maxybleach.0.summ <- summary(maxybleach.0.lsm)
-
-# look at slopes and 100% clade D Fv/Fm
-maxybleach.1.lsm <- lsmeans(maxybleach.m, "Species", 
-                          at=list(PropD.bleach=1, BleachLvl="Med"))  
-cld(maxybleach.1.lsm) # View lsmean comparisons
-plot(maxybleach.1.lsm, main="Bleached Fv/Fm with 100% clade D", horizontal=F)
-maxybleach.1.summ <- summary(maxybleach.1.lsm)
-
-# Test differences in slopes
-maxybleach.lst <- lstrends(maxybleach.m, specs="Species", var="PropD.bleach")
-cld(maxybleach.lst) # Slopes of fit for each species
-#plot(Effect(c("PropD.bleach", "Species"), maxybleach.m))  # Plot model fit
-```
-
-* With no clade D, *O.
-faveolata* maintained significantly higher Fv/Fm (`r maxybleach.0.summ$lsmean[1]`±`r maxybleach.0.summ$SE[1]`) than *S.
-siderea*, while *M.
-cavernosa* was also lower but not significantly different.
-* *O. faveolata* and *S. siderea* with 100% clade D had the same Fv/Fm (0.31) under heat stress, while *M. cavernosa* with clade D had a significantly higher Fv/Fm (0.38). 
-* As the proportion of clade D went down, Fv/Fm went down > 3 times faster in *S.
-siderea* and almost 5 times faster in *M.
-cavernosa* compa#d95f02 to *O.
-faveolata*, suggesting their other symbionts differ in thermal sensitivity.
-
-```{r, eval=T}
 # Melt data frame so that "time" (i.e., initial vs. bleached) can be used as factor
@@
 bleachdat$time <- factor(bleachdat$time)
+```
 
-# Test of slopes of FvFm ~ PropD are different for different BleachLvls
+# Analyze photochemical efficiency (Fv/Fm)
+```{r analyze_fvfm}
+# Test of slopes of bleached FvFm ~ PropD are different for different BleachLvls
 testmod <- lmer(MaxY ~ PropD * Species * BleachLvl + (1|Colony), 
@@
 
+## Plot Figure 1
 ```{r Fig1}
@@
 ```
-![Figure 1.](../figures/Fig1.png)
+![Figure 1.
+Photochemical efficiency as a function of the proportion of clade D *Symbiodinium*.
+Solid lines represent ambient conditions, and dotted lines represent the relationship under heat stress (32°C).
+Shaded regions represent 90% confidence intervals around fitted values.](../figures/Fig1.png)
 
@@
 ```
-![Figure 2](../figures/Fig2.png)
+![Figure 2.
+Photochemical advantage of clade D symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars).
+Values indicate the difference in Fv/Fm of corals with all and no clade D symbionts (i.e., slopes of the lines in Fig.
+1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)
 
 # Symbiont shuffling after bleaching
-```{r}
+```{r analyze_symshuff}
 recovdat <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "Total.init", "PropD.init", "MaxY.init", "BleachLvl", "RecovTemp", "Total.recov", "PropD.recov", "MaxY.recov")]
@@
 summary(propDmod)  # Model summary
-dropterm(propDmod, test="F")  # Partial F-tests: Results in Table 1
+dropterm(propDmod, test="F")  # Partial F-tests
 
@@
 ```
-![Figure 3.](../figures/Fig3.png)
+![Figure 3.
+Symbiont shuffling in three coral species.
+Panel **(A)** shows the proportion of clade D *Symbiodinium* before bleaching and after recovery, with the diagonal line of identity representing no change in symbiont proportions.
+Fitted values for each species (indicated by color) and heat stress duration (indicated by line type) are plotted with shaded areas representing 95% confidence intervals.
+Panel **(B)** shows the integrated symbiont shuffling response (from before bleaching to after recovery) for each species and bleaching treatment, calculated as the area above or below the diagonal line of identity, normalized to a range of -1 (i.e., a complete loss of clade D symbionts) to +1 (i.e., complete dominance by clade D symbionts).](../figures/Fig3.png)
 
@@
                      type="response")
+with(summary(glmod2), 1-(deviance(glmod2)/null.deviance))
+
 plot(df$shuffle ~ df$ADH.days, ylim=c(-1,1), yaxs="r", xpd=NA,
@@
 ```
-![Figure 4.](../figures/Fig4.png)
-
-# Analyze data using initial *M. cavernosa* bleaching experiment
-# Analyze photochemical efficiency (Fv/Fm)
-
-### Initial Fv/Fm before stress
-```{r MaxY.init, fig.width=7, fig.height=4}
-
-# Plot Fv/Fm vs. PropD by Species
-xyplot(MaxY.init ~ PropD.init | Species, alldata, 
-       main="Initial Fv/Fm by species",
-       xlab="Initial proportion clade D", ylab="Initial Fv/Fm")
-
-# Fit mixed model with fixed effects of Species and proportion clade D
-maxyinit.m <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
-                   data=alldata) 
-anova(maxyinit.m)  # Assess significance of Species
-
-# Calculate mean Fv/Fm for each species at PropD=0 w/ pairwise comparisons
-maxyinit.lsm <- lsmeans(maxyinit.m, "Species", at=list(PropD.init=0))  
-cld(maxyinit.lsm) # View lsmean comparisons
-plot(maxyinit.lsm, main="Initial Fv/Fm with no clade D", horizontal=F)
-
-# Effect of proportion clade D on initial Fv/Fm
-## (Excludes *M. cavernosa*, which did not host clade D before stress)
-
-# Look at distribution of PropD within each colony
-#histogram(~ PropD.init | Colony, data=subset(alldata, Species!="Mc"))
-# Not enough variation in PropD within colonies to be able to use random slopes
-
-# Fit mixed model with fixed effect of Species and PropD, random effect of Colony
-maxyinit.m2 <- lmer(MaxY.init ~ PropD.init * Species + (1|Colony),
-                    data=subset(alldata, Species!="Mc"))
-anova(maxyinit.m2)  # Assess significance of species and propD interaction
-maxyinit.m2.lst <- lstrends(maxyinit.m2, specs="Species", var="PropD.init")
-cld(maxyinit.m2.lst) # Slopes of fit for each species
-plot(Effect(c("PropD.init", "Species"), maxyinit.m2))  # Plot model fit
-
-# Difference in Fv/Fm between Ofav and Ssid with 100% clade D?
-maxyinit.lsm <- lsmeans(maxyinit.m2, "Species", at=list(PropD.init=1))
-cld(maxyinit.lsm)
-plot(maxyinit.lsm, main="Initial Fv/Fm with 100% clade D", horizontal=F)
-```
-
-* Higher amounts of clade D #d95f02uce Fv/Fm, with this effect being stronger in *O.
-faveolata* compa#d95f02 to *S.
-siderea*.
-However, Fv/Fm with 100% D is similar in both species, indicating that it is the higher Fv/Fm of the non-clade-D symbiont of *O.
-faveolata* that is driving this trend.
-
-### Fv/Fm after heat stress
-```{r, fig.width=7, fig.height=4}
-
-# Subset corals bleached by heat
-## Corals bleached by heat for 10 days only
-heat10day <- subset(alldata, BleachLvl=="Med" | BleachLvl=="Heat")
-## All corals bleached by heat (7, 10, and 14 days for Ofav and Ssid)
-heatbleach <- subset(alldata, !BleachLvl %in% c("DCMU", "Control"))
-## Set BleachLvl for heated Mcav to "Med" (=10 days of heat) to match Ofav and Ssid
-heatbleach[heatbleach$Species=="Mc", "BleachLvl"] <- "Med"
-
-# Plot bleached Fv/Fm vs. bleached proportion clade D by species
-xyplot(MaxY.bleach ~ PropD.bleach | Species, data=heatbleach)
-
-# Look at distribution of PropD after bleaching within each colony
-# histogram(~ PropD.bleach | Colony, data=heatbleach)
-## Many colonies do not have a good spread of proportion clade D.
-Therefore, allowing random slopes for each colony when modeling responses as a function of proportion clade D will give very high variance (i.e., there is not enough data to constrain slopes within colonies).
-Therefore, we will model responses with random intercepts only.
-
-# Fit mixed model with fixed effects of Species and proportion clade D and duration
-maxybleach.m <- lmer(MaxY.bleach ~ PropD.bleach * Species * BleachLvl + (1|Colony),
-                     data=heatbleach)
-select <- lmerTest::step(maxybleach.m)$model
-maxybleach.m <- lmer(formula(select), data=heatbleach)
-anova(maxybleach.m)
-## Not there is NO significant interaction between PropD and BleachLvl, indicating that the duration of heat stress did not influence the relationship between PropD and Fv/Fm.
-Therefore, we can use data from all three BleachLvl's (for Ofav and Ssid) to compare the effect of PropD on Fv/Fm among species.
-
-# Calculate mean Fv/Fm for each species at PropD=0 and BleachLvl=Med
-maxybleach.0.lsm <- lsmeans(maxybleach.m, "Species", 
-                          at=list(PropD.bleach=0, BleachLvl="Med"))  
-cld(maxybleach.0.lsm) # View lsmean comparisons
-plot(maxybleach.0.lsm, main="Bleached Fv/Fm with 0% clade D", horizontal=F)
-maxybleach.0.summ <- summary(maxybleach.0.lsm)
-
-# Re-fit model excluding Mcav to look at slopes and 100% clade D Fv/Fm
-maxybleach.m.noMc <- update(maxybleach.m, data=subset(heatbleach, Species!="Mc"))
-
-maxybleach.1.lsm <- lsmeans(maxybleach.m.noMc, "Species", 
-                          at=list(PropD.bleach=1, BleachLvl="Med"))  
-cld(maxybleach.1.lsm) # View lsmean comparisons
-plot(maxybleach.1.lsm, main="Bleached Fv/Fm with 100% clade D", horizontal=F)
-maxybleach.1.summ <- summary(maxybleach.1.lsm)
-
-# Test differences in slopes
-maxybleach.lst <- lstrends(maxybleach.m.noMc, specs="Species", var="PropD.bleach")
-cld(maxybleach.lst) # Slopes of fit for each species
-plot(Effect(c("PropD.bleach", "Species"), maxybleach.m.noMc))  # Plot model fit
-```
-
-* With no clade D, *O. faveolata* maintained significantly higher Fv/Fm (`r maxybleach.0.summ$lsmean[1]`±`r maxybleach.0.summ$SE[1]`) than the other two species under heat stress.
-* *O.
-faveolata* and *S.
-siderea* with 100% clade D had the same Fv/Fm (0.31) under heat stress, suggesting that clade D performs the same in both host species under stress.
-As the proportion of clade D went down, Fv/Fm went down > 3 times faster in *S.
-siderea* compa#d95f02 to *O.
-faveolata*, suggesting their other symbionts differ in thermal sensitivity (*M.
-cavernosa* still did not have clade D at this point).
+![Figure 4.
+Symbiont shuffling as a function of the photochemical advantage of clade D and heat stress duration.
+Panel **(A)** shows the symbiont shuffling response (as in Fig.
+3B) as a function of heat stress duration (days) and the photochemical advantage of clade D under heat stress (A^D^~Heat~), with lines corresponding to the A^D^~Heat~ values for the three coral species studied (indicated by color), and two hypothetical intermediate values (gray lines).
+Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (psuedo-R^2^=0.93) by the time-integrated photochemical advantage of clade D under heat stress.](../figures/Fig4.png)

### 2017-06-16 analysis/analysis.Rmd
reorganize file structure

--- 2017-06-16 supp/supp.Rmd
@@
                      type="response")
-with(summary(glmod2), 1-(deviance(glmod2)/null.deviance))
+pseudoR2 <- with(summary(glmod2), 1-(deviance(glmod2)/null.deviance))
 
@@
 Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (psuedo-R^2^=0.93) by the time-integrated photochemical advantage of clade D under heat stress.](../figures/Fig4.png)
+
+```{r save_workspace}
+save.image(file="analysis/analysis.RData")
+```

### 2017-06-16 R/Fig1.R

@@ -1,49 +1,0 @@
-library(lme4)
-library(reshape2)
-library(effects)
-library(lsmeans)
-
-# Load data -----
-rm(list=ls())
-alldata <- read.csv("data/master.csv")
-alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
-alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
-
-# Init Fv/Fm for Mcav during second heat bleaching (=same as recov from 1st)
-Mc <- subset(alldata, Species=="Mc" & BleachLvl2=="Heat")
-#plot(MaxY.recov ~ PropD.recov, Mc)
-mcmaxyrecov.m <- lmer(MaxY.recov ~ PropD.recov + (PropD.recov|Colony), Mc)
-anova(mcmaxyrecov.m)
-#plot(Effect("PropD.recov", mcmaxyrecov.m))
-#plot(MaxY.bleach2 ~ PropD.bleach2, Mc)
-mcmaxybleach2.m <- lmer(MaxY.bleach2 ~ PropD.bleach2 + (PropD.recov|Colony), Mc)
-#plot(Effect("PropD.bleach2", mcmaxybleach2.m))
-
-# How much is Fv/Fm reduced in Mcav because they were previously bleached? - can't tell because changed to D - confounded
-
-# Get data frame with Of, Ss 10-day bleach  (Med bleach only or all bleachlvls pooled?)
-#SsOf <- subset(alldata, BleachLvl=="Med" & Species!="Mc")
-SsOf <- subset(alldata, Species!="Mc")
-# Get data frame with Mc 10-day bleach (from repeat bleaching experiment)
-Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
-Mc$PropD.init <- Mc$PropD.recov  # Recov from initial bleach = init for repeat bleach / for merging with Ss, Of
-Mc$MaxY.init <- Mc$MaxY.recov
-Mc$PropD.bleach <- Mc$PropD.bleach2  # repeat bleach = bleach / for merging with Ss, Of
-Mc$MaxY.bleach <- Mc$MaxY.bleach2
-# Merge three species together (from two separate but comparable bleaching experiments)
-bleach10day <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "PropD.init", "MaxY.init", "PropD.bleach", "MaxY.bleach")]
[…]

### 2017-06-16 R/Fig2.R

@@ -1,87 +1,0 @@
-
-library(lme4)
-library(reshape2)
-library(effects)
-library(lsmeans)
-
-# Load data -----
-rm(list=ls())
-alldata <- read.csv("data/master.csv")
-alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
-alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))
-
-
-
-
-
-
-
-# Init Fv/Fm for Mcav during second heat bleaching (=same as recov from 1st)
-Mc <- subset(alldata, Species=="Mc" & BleachLvl2=="Heat")
-#plot(MaxY.recov ~ PropD.recov, Mc)
-mcmaxyrecov.m <- lmer(MaxY.recov ~ PropD.recov + (PropD.recov|Colony), Mc)
-anova(mcmaxyrecov.m)
-#plot(Effect("PropD.recov", mcmaxyrecov.m))
-#plot(MaxY.bleach2 ~ PropD.bleach2, Mc)
-mcmaxybleach2.m <- lmer(MaxY.bleach2 ~ PropD.bleach2 + (PropD.recov|Colony), Mc)
-#plot(Effect("PropD.bleach2", mcmaxybleach2.m))
-
-# How much is Fv/Fm reduced in Mcav because they were previously bleached? - can't tell because changed to D - confounded
-
-# Get data frame with Of, Ss 10-day bleach  (Med bleach only or all bleachlvls pooled?)
-#SsOf <- subset(alldata, BleachLvl=="Med" & Species!="Mc")
-SsOf <- subset(alldata, Species!="Mc")
-# Get data frame with Mc 10-day bleach (from repeat bleaching experiment)
-Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
-Mc$PropD.init <- Mc$PropD.recov  # Recov from initial bleach = init for repeat bleach / for merging with Ss, Of
-Mc$MaxY.init <- Mc$MaxY.recov
-Mc$PropD.bleach <- Mc$PropD.bleach2  # repeat bleach = bleach / for merging with Ss, Of
-Mc$MaxY.bleach <- Mc$MaxY.bleach2
-# Merge three species together (from two separate but comparable bleaching experiments)
-bleach10day <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "PropD.init", "MaxY.init", "PropD.bleach", "MaxY.bleach")]
[…]

### 2017-06-16 manuscript/manuscript.Rmd
write methods section

@@
 ---
-title: Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
+title: Symbiont shuffling linked to differential photochemical dynamics of *Symbiodinium* in three Caribbean reef corals
 author: "Ross Cunning"
@@
 
-Title ideas:
+```{r load_analysis}
+load("../analysis/analysis.RData")
 
-Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
+ncolsperspp <- with(bleach3c, aggregate(Colony ~ Species, 
+                                        FUN=function(x) length(unique(x))))
+nOfavcol <- ncolsperspp[1, 2]
+nSsidcol <- ncolsperspp[2, 2]
+nMcavcol <- ncolsperspp[3, 2]
+
+ncorepercol <- with(bleach3c, table(Species, Colony))
+ncorepercol[ncorepercol==0] <- NA
+rangeOfav <- range(ncorepercol[1, ], na.rm=T)
+rangeSsid <- range(ncorepercol[2, ], na.rm=T)
+rangeMcav <- range(ncorepercol[3, ], na.rm=T)
+
+ncoreperspp <- with(bleach3c, table(Species))
+nOfavcore <- ncoreperspp["Of"]
+nSsidcore <- ncoreperspp["Ss"]
+nMcavcore <- ncoreperspp["Mc"]
+
+```
+
+# Abstract
+
+# Introduction
+
+# Materials and Methods
+
+### Experimental manipulations
+
+Two separate bleaching and recovery experiments were conducted.
+The first experiment included both *O.
+faveolata* (`r nOfavcol` colonies, `r min(rangeOfav)`-`r max(rangeOfav)` cores colony^-1^, n=`r nOfavcore` cores) and *S.
+siderea* (`r nSsidcol` colonies, `r min(rangeSsid)`-`r max(rangeSsid)` cores colony^-1^, n=`r nSsidcore` cores), with bleaching treaments of exposure to 32°C for 7, 10, and 14 days [@Cunning:2015ja], and recovery for 3 months at 24 or 29°C.
+The second experiment was conducted on *M.
+cavernosa* (`r nMcavcol` colonies, `r min(rangeMcav)`-`r max(rangeMcav)` cores colony^-1^, n=`r nMcavcore` cores), with a bleaching treatment of 32°C for 10 days [@Silverstein:2015er] and recovery for 3 months at 24 or 29°C.
+Prior to experimentation, the *O.
+faveolata* cores contained varying mixtures of *Symbiodinium* in clades B and D [@Cunning:2015ja], while *S.
+siderea* colonies contained mixtures of clades C and D [@Cunning:2013tl].
+*M.
+cavernosa* contained only clade C *Symbiodinium* upon collection, but an initial bleaching and recovery experiment involving some of these cores produced a range of mixtures of clades C and D [@Silverstein:2015er].
+Both experiments were conducted in the same indoor, semi-recirculating coral culture facility at the University of Miami.
+For additional details on coral collection and experimental set-up, see @Silverstein:2015er and @Cunning:2015ja.
+
+### Sampling and data collection
+
+In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of the heat stress treatment using an Imaging PAM (Walz, Effeltrich, Germany).
+At the same time points, small biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR.
+Tissue samples were collected again after 3 months of recovery from bleaching to analyze changes in *Symbiodinium* community structure (i.e., symbiont shuffling).
+For additional details on sample collection and processing, see @Silverstein:2015er and @Cunning:2015ja.
+
+### Data analysis
+
+Photochemical efficiency (*F~v~/F~m~*) was analyzed by a linear mixed model (*lmer*; [@Bates:2015wh]) with fixed effects of time (pre- or post-heat-stress), proportion clade D (measured at the same time point), and coral species, and with random intercepts for each colony.
+Backward elimination of non-significant effects from a fully crossed model was performed using partial F-tests in *lmerTest* [@lmerTestTestsinL:j0go4ZGL].
+Regression coefficients for the proportion clade D were interpreted as the photochemical advantage of clade D for each species under ambient (pre-heat-stress; A^D^~Amb~) and heated (post-heat-stress; A^D^~Heat~) conditions, and tested for statistical differences by pairwise comparisons in *lsmeans* [@Lenth:2015vl].
+Days of heat stress was not included as a predictor of *F~v~/F~m~*, since analysis of corals in the first experiment indicated it had no impact on the photochemical advantage of clade D (i.e., no significant interaction between stress duration and proportion clade D).
+
+Symbiont shuffling was analyzed by modeling the final proportion of clade D (after 3 months of recovery) as a function of the initial proportion clade D (before bleaching), using a quasi-binomial generalized linear model (GLM) with species, bleaching duration, and recovery temperature as co-predictors.
+Fitted values for the final proportion of clade D for each species at each bleaching duration (averaged across recovery temperatures, and predicted for *M.
+cavernosa* at 7- and 14-day durations using the *effects* package [@Fox:2003uu]), were plotted against initial proportion of clade D with 90% confidence intervals.
+From these fitted values, an integrated metric of symbiont shuffling was generated across the full range of initial clade D proportions by finding the area between the fitted curve for each group and the 45° line of identity, which indicates no change in symbiont proportions.
+This metric was scaled to a range of -1 (e.g., complete loss of clade D) to +1 (e.g., complete dominance by clade D), with zero indicating no symbiont shuffling.
+
+Finally, a link between the differential performance of symbionts and symbiont shuffling was investigated by modeling (with a quasi-binomial GLM) the integrated symbiont shuffling metric as a function of the photochemical advantage of clade D under heat stress (A^D^~Heat~) for each species, and days of exposure to heat stress.
+All data analysis was conducted using R v3.3.1 [@RALanguageandEn:2014wf].
+Data and R code to reproduce all analyses and figures presented here are available at: github.com/jrcunning/symshuff-3c.
+
+# Results
+
+# Discussion
+
+# Acknowledgements
+
+# References

### 2017-06-19 analysis/analysis.Rmd
add stats to Fig2

@@
 bleachmod <- lmer(formula(select), data=bleachdat)
-anova(bleachmod)
+lmerTest::anova(bleachmod)
 
 plot(Effect(c("PropD", "time", "Species"), bleachmod), multiline=T, ci.style="bands")
-
-# Compare slopes of Fv/Fm vs. PropD among species by time
-lst <- lstrends(bleachmod, specs=c("Species", "time"), var="PropD")
-cld(contrast(lst, by="time"))
-plot(lst, horizontal=F)
 
@@
 ```{r Fig2}
+
+# Compare slopes of Fv/Fm vs. PropD among species by time
+lst <- lstrends(bleachmod, specs=c("Species", "time"), var="PropD")
+lst.pw <- summary(contrast(lst, "pairwise", by="time"))
+l <- cld(contrast(lst), Letters=c("abcdefg"))
+grps <- matrix(t(matrix(l$.group, ncol=2)))
+grps <- stringr::str_trim(grps)
+
 # Plot Advantage of D vs. other symbiont in each coral under
@@
       line=1.5, cex=0.8)
+l.pos <- ifelse(sign(matrix(ADf))==-1,
+                matrix(ADl)-.01, matrix(ADu)+.01)
+text(bars, l.pos, grps, xpd=NA, cex=0.6)
 #legend("topleft", pch=22, pt.bg=c("gray80", "gray20"), y.intersp=1.2,
@@
 # Get fitted values and plot shuffling ~ days for different relative advantages of D
-stresslevs <- c(0.033, 0.05, 0.07, 0.103, 0.163)
-cols <- c("#1b9e77", "gray40", "gray40", "#7570b3", "#d95f02")
+stresslevs <- c(0.01, 0.033, 0.05, 0.07, 0.103, 0.163)
+cols <- c("gray40", "#1b9e77", "gray40", "gray40", "#7570b3", "#d95f02")
 resdf <- expand.grid(days=7:14, ADH=stresslevs)

### 2017-06-19 analysis/analysis.Rmd
fix Fig4 colors

@@
      ylab="", mgp=c(1.8,0.1,0),
-     col=c("#d95f02", "#1b9e77", "#7570b3")[as.numeric(factor(df$Species))])
+     col=c("#1b9e77", "#7570b3", "#d95f02")[as.numeric(factor(df$Species))])
 mtext("Symbiont shuffling", side=2, line=1.5)

### 2017-07-13 manuscript/manuscript.Rmd
draft

@@
 title: Symbiont shuffling linked to differential photochemical dynamics of *Symbiodinium* in three Caribbean reef corals
-author: "Ross Cunning"
-date: "6/7/2017"
-output: html_document
+author:
+  - name: Ross Cunning
+    email: ross.cunning@gmail.com
+    affiliation: MBE
+    footnote: Corresponding Author
+  - name: Andrew C. Baker
+    affiliation: MBE
+address:
+  - code: MBE
+    address: Department of Marine Biology and Ecology, Rosenstiel School of Marine and Atmospheric Science, University of Miami, Miami FL, USA
+bibliography: library.bib
+csl: coral-reefs.csl
+output:
+  word_document:
+  rticles::elsevier_article:
+    fig_caption: yes
+
 ---
+```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
+knitr::opts_knit$set(root.dir = "..")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(lmerTest)
+library(MASS)
+```
 
 ```{r load_analysis}
-load("../analysis/analysis.RData")
+load("analysis/analysis.RData")
 
@@
 # Abstract
+  Dynamic symbioses with functionally diverse dinoflagellate algae in the genus *Symbiodinium* may allow some reef corals to alter their phenotypes through 'symbiont shuffling', or changes in symbiont community composition.
+In particular, corals have been shown to become more bleaching-resistant by increasing the relative abundance of thermally tolerant *Symbiodinium* in clade D after bleaching.
+Despite the immediate relevance of this phenomenon to corals living in warming oceans -- and to interventions aimed at boosting coral resilience -- the mechanisms governing how, why, and when symbiont shuffling occurs are still poorly understood.
+Here, we performed controlled thermal bleaching and recovery experiments on three species of Caribbean reef-building corals, and show that the degree of symbiont shuffling is determined by (1) the difference in photochemical efficiency (*F~v~/F~m~*) of co-occurring symbionts under stress, and (2) the duration of stress exposure.
+The photochemical advantage of clade D under stress was greatest in *Montastraea cavernosa*, intermediate in *Siderastrea siderea*, and lowest in *Orbicella faveolata*, and correlated positively with the magnitude of increase in clade D.
+In holobionts where clade D had less of an advantage (i.e.
+a smaller difference in *F~v~/F~m~*), longer durations of heat stress were required to elicit commensurate increases in clade D.
+These findings help elucidate general rules governing the process of symbiont shuffling in reef corals, which will help refine predictions of coral responses to environmental change, and inform interventions to manipulate symbiont communities to enhance coral resilience.
 
@@
 
+Coral reef ecosystems are declining globally at an alarming rate, owing primarily to mass coral bleaching caused by anthropogenic increases in sea surface temperature [@Hughes:2017bs].
+Heat-induced bleaching, or the loss of corals' nutritional symbionts (*Symbiodinium* spp.; @Jokiel:1977p7353), often leads to coral mortality, which precipitates reef ecosystem degradation [@Baker:2008p3421].
+
+However, depending on the severity of bleaching, some corals may recover their *Symbiodinium* communities within weeks to months [@Jokiel:1977p7353, @Jones:1997p3499, @Cunning:2016jb], and ultimately return to a healthy state.
+Moreover, the loss and subsequent re-establishment of the symbiotic algal community may provide an opportunity for change in its composition, i.e., the relative abundance of different *Symbiodinium* types [@Buddemeier:1993p3, @Baker:2003p200].
+When such 'symbiont shuffling' occurs, the new configuration of the *Symbiodinium* assemblage may be better suited to tolerate future bleaching stressors [@Berkelmans:2006p191, @Silverstein:2015er], representing a mechanism of rapid ecological acclimatization for corals.
+Moreover, directed manipulation of coral symbiont communities has been proposed as a mechanism to increase the resilience of reef corals.
+
+*Symbiodinium* is a highly diverse genus comprised of nine clades [@Pochon:2014bk], which are further subdivided into numerous types (with species descriptions underway [@LaJeunesse:2014io, @Wham:2017dy]) with varying functional biology [@Suggett:2015dv].
+*S.
+trenchii*, a member of *Symbiodinium* clade D, is a heat-tolerant symbiont that confers a bleaching threshold ~1-2°C higher than other *Symbiodinium* types [@Silverstein:2017ep].
+Increases in the abundance of clade D symbionts have been observed during and after bleaching in several coral species [@Jones:2008p113, @LaJeunesse:2009p7364], and increases in clade A have been observed in others [@Grottoli:2014jo].
+Many more corals have been observed in association with *Symbiodinium* from multiple clades [@Silverstein:2012ub], suggesting there is potential for symbiont shuffling to occur.
+
+Nevertheless, the mechanisms governing the process of symbiont shuffling, and its prevalence among corals, are still poorly understood.
+In *Orbicella faveolata*, symbiont shuffling was shown to depend on the severity of bleaching (i.e., the magnitude of disturbance to the existing community), as well as the environmental conditions during recovery [@Cunning:2015ja], suggesting that the relative performance of different symbionts influences the trajectory of symbiont community re-establishment.
+However, whether general rules about symbiont performance can explain variability in symbiont shuffling among coral species with different *Symbiodinium* types has not been investigated.
+Ultimately, a deeper understanding of symbiont community dynamics in corals is a critical foundation for 1) predicting corals responses to environmental change [@Logan:2013hp], and 2) manipulating symbiont communities as a way to increase resilience in coral reef restoration [@vanOppen:2015es].
+
+Here, we investigate symbiont shuffling through controlled bleaching experiments with three Caribbean coral species hosting mixed *Symbiodinium* assemblages: *O.
+faveolata* (with clades B and D), *S.
+siderea* (with clades C and D), and *M.
+cavernosa* (with clades C and D).
+We used chlorophyll fluorescence in combination with clade-specific quantitative PCR assays to explore links between the photochemical efficiencies of co-occurring symbionts and the magnitude of symbiont shuffling observed after bleaching.
+
 # Materials and Methods
 
-### Experimental manipulations
+## Experimental manipulations
 
@@
 
-### Sampling and data collection
+## Sampling and data collection
 
-In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of the heat stress treatment using an Imaging PAM (Walz, Effeltrich, Germany).
-At the same time points, small biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR.
+In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of heat stress using an Imaging PAM (Walz, Effeltrich, Germany).
+At the same time points, small tissue biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR.
 Tissue samples were collected again after 3 months of recovery from bleaching to analyze changes in *Symbiodinium* community structure (i.e., symbiont shuffling).
-For additional details on sample collection and processing, see @Silverstein:2015er and @Cunning:2015ja.
+For additional details on sampling and data collection, see @Silverstein:2015er and @Cunning:2015ja.
 
-### Data analysis
+## Data analysis
 
@@
 Backward elimination of non-significant effects from a fully crossed model was performed using partial F-tests in *lmerTest* [@lmerTestTestsinL:j0go4ZGL].
-Regression coefficients for the proportion clade D were interpreted as the photochemical advantage of clade D for each species under ambient (pre-heat-stress; A^D^~Amb~) and heated (post-heat-stress; A^D^~Heat~) conditions, and tested for statistical differences by pairwise comparisons in *lsmeans* [@Lenth:2015vl].
-Days of heat stress was not included as a predictor of *F~v~/F~m~*, since analysis of corals in the first experiment indicated it had no impact on the photochemical advantage of clade D (i.e., no significant interaction between stress duration and proportion clade D).
+By relating the photochemical efficiency of *Symbiodinium* communities to their taxonomic composition [*sensu*  @Suggett:2009p4201], we can tease apart the relative photochemical performance of the two different symbionts in each coral.
+Specifically, regression coefficients for the proportion clade D were interpreted as the photochemical advantage of clade D for each species under ambient (pre-heat-stress; *A^D^~Amb~*) and heated (post-heat-stress; *A^D^~Heat~*) conditions, and tested for statistical differences by pairwise comparisons using *lsmeans* [@Lenth:2015vl].
+Stress duration (i.e., 7, 10, or 14 days) was not included as a predictor since it had no impact on the photochemical advantage of clade D in the first experiment (i.e., no significant interaction between stress duration and proportion clade D).
 
-Symbiont shuffling was analyzed by modeling the final proportion of clade D (after 3 months of recovery) as a function of the initial proportion clade D (before bleaching), using a quasi-binomial generalized linear model (GLM) with species, bleaching duration, and recovery temperature as co-predictors.
-Fitted values for the final proportion of clade D for each species at each bleaching duration (averaged across recovery temperatures, and predicted for *M.
+Symbiont shuffling was analyzed by modeling the final proportion of clade D (after 3 months of recovery) as a function of the initial proportion clade D (before bleaching), using a quasi-binomial generalized linear model (GLM) with species, stress duration, and recovery temperature as co-predictors.
+Fitted values for the final proportion of clade D for each species at each stress duration (averaged across recovery temperatures, and predicted for *M.
 cavernosa* at 7- and 14-day durations using the *effects* package [@Fox:2003uu]), were plotted against initial proportion of clade D with 90% confidence intervals.
-From these fitted values, an integrated metric of symbiont shuffling was generated across the full range of initial clade D proportions by finding the area between the fitted curve for each group and the 45° line of identity, which indicates no change in symbiont proportions.
+From these fitted values, an integrated metric of symbiont shuffling was generated across the full range of initial clade D proportions by finding the area between the fitted curve for each group and the diagonal line of identity, which indicates no change in symbiont proportions.
 This metric was scaled to a range of -1 (e.g., complete loss of clade D) to +1 (e.g., complete dominance by clade D), with zero indicating no symbiont shuffling.
 
-Finally, a link between the differential performance of symbionts and symbiont shuffling was investigated by modeling (with a quasi-binomial GLM) the integrated symbiont shuffling metric as a function of the photochemical advantage of clade D under heat stress (A^D^~Heat~) for each species, and days of exposure to heat stress.
+Finally, we used a quasi-binomial GLM to model the integrated symbiont shuffling response as a function of the photochemical advantage of clade D under heat stress (A^D^~Heat~) for each species, and days of exposure to heat stress.
 All data analysis was conducted using R v3.3.1 [@RALanguageandEn:2014wf].
@@
 
+## Photochemistry of symbiont assemblages
+
+Measured values of *F~v~/F~m~* were significantly affected by the proportion of clade D in symbiont assemblages (Fig.
+1), but this effect varied among coral species and between ambient and heat-stress conditions (three-way interaction, *F*=`r round(lmerTest::anova(bleachmod)$F[7], 3)`, *p*=`r round(lmerTest::anova(bleachmod)$P[7], 3)`).
+In all three coral species, *F~v~/F~m~* was negatively related to the proportion of clade D under ambient conditions, but positively related to the proportion of clade D under heat stress.
+The slope of each line in Fig.
+1 (i.e., the difference in *F~v~/F~m~* when proportion D=1 and when proportion D=0) can be interpreted as the photochemical advantage of clade D compared to the other symbiont (Fig.
+2).
+The advantage of D under ambient conditions (*A^D^~Amb~*) was significantly more negative (i.e., a greater disadvantage) in *O.
+faveolata* (`r round(AD[4,3], 3)`) compared to *M.
+cavernosa* (`r round(AD[6,3], 3)`, *p*=`r round(lst.pw$p.value[5], 3)`), and intermediate (`r round(AD[5,3], 3)`, but not significantly different, *p*>`r round(min(lst.pw$p.value[c(4,6)]), digits=1)`) in *S.
+siderea*.
+Under heat stress, the advantage of clade D (*A^D^~Heat~*) was greatest in *M.
+cavernosa* (`r round(AD[3,3], 3)`), intermediate in *S.
+siderea* (`r round(AD[2,3], 3)`), and lowest in *O.
+faveolata* (`r round(AD[1,3], 3)`, all pairwise comparisons, *p*<`r plyr::round_any(max(lst.pw$p.value[c(1,2,3)]), 0.001, ceiling)`).
+
+![Figure 1.
+Photochemical efficiency as a function of the proportion of clade D *Symbiodinium* in **A)** *O.
+faveolata*, **B)** *S.
+siderea*, and **C)** *M.
+cavernosa*.
+Solid lines represent ambient conditions, and dotted lines represent the relationship under heat stress (32$^\circ$C).
+Shaded regions represent 90% confidence intervals around fitted values.](../figures/Fig1.png)
+
+![Figure 2.
+Photochemical advantage of clade D symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars).
+Values indicate the difference in Fv/Fm of corals with all and no clade D symbionts (i.e., slopes of the lines in Fig.
+1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)
+
+## Symbiont shuffling
+
+The proportion of clade D after recovery from bleaching (Fig.
+3A) depended on the initial proportion of clade D (*F*=`r round(dropterm(propDmod, test="F")$F[2], 3)`, *p*`r format.pval(dropterm(propDmod, test="F")$P[2], eps=0.0001)`) and an interaction among coral species, heat stress duration, and recovery temperature (*F*=`r round(dropterm(propDmod, test="F")$F[3], 3)`, *p*=`r round(dropterm(propDmod, test="F")$P[3], 3)`).
+Fitted values were averaged across recovery temperatures in order to focus on differences among species and heat stress durations.
+Across the full range of initial clade D proportions, the integrated symbiont shuffling response, $S$ (Fig.
+3B), showed a loss of clade D in *O.
+faveolata* after recovery from 7 and 10 days of heat stress ($S$=`r round(auc[4,3], 3)` and `r round(auc[7,3], 3)`), but a gain of clade D after recovery from 14 days of heat stress ($S$=`r round(auc[1,3], 3)`).
+In contrast, *S.
+siderea* underwent greater shuffling toward clade D after 7 ($S$=`r round(auc[5,3], 3)`), 10 ($S$=`r round(auc[8,3], 3)`), and 14 ($S$=`r round(auc[2,3], 3)`) days of heat stress.
+*M.
+cavernosa* shuffled almost completely to clade D assemblages following 10 days of stress ($S$=`r round(auc[9,3], 3)`), and thus the statistical model also predicted near complete shuffling to D after 7 ($S$=`r round(auc[6,3], 3)`) and 14 (($S$=`r round(auc[3,3], 3)`) days of heat stress.
+
+![Figure 3.
+Symbiont shuffling in three coral species.
+Panel **(A)** shows the proportion of clade D *Symbiodinium* before bleaching and after recovery, with the diagonal line of identity representing no change in symbiont proportions.
+Fitted values for each species (indicated by color) and heat stress duration (indicated by line type) are plotted with shaded areas representing 95% confidence intervals.
+Panel **(B)** shows the integrated symbiont shuffling response (from before bleaching to after recovery) for each species and bleaching treatment, calculated as the area above or below the diagonal line of identity, normalized to a range of -1 (i.e., a complete loss of clade D symbionts) to +1 (i.e., complete dominance by clade D symbionts).](../figures/Fig3.png)
+
+## Relationship between photochemistry and symbiont shuffling
+
+The magnitude of symbiont shuffling showed a significant statistical relationship (Fig.
+4A) with both the photochemical advantage of clade D under heat stress (*A^D^~Heat~*, *F*=`r round(anova(glmod, test="F")$F[2], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[2], eps=0.0001)`) and the duration of heat stress (*F*=`r round(anova(glmod, test="F")$F[3], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[3], eps=0.0001)`).
+Specifically, increases in the amount of clade D were greater for higher values of *A^D^~Heat~* and longer durations of heat stress.
+The model predicts the duration of heat stress required to cause symbiont shuffling toward clade D for different values of *A^D^~Heat~* (Fig.
+4A), and shows that 7 days of heat stress results in shuffling toward clade D for *A^D^~Heat~*>~0.06.
+Lower values of *A^D^~Heat~* cause a decrease in the proportion of clade D after recovery, even following 14 days of heat stress for *A^D^~Heat~*<~0.01.
+
+Across these three coral species, the interaction of *A^D^~Heat~* and the duration of heat stress explains `r round(pseudoR2*100, 1)`% of the variability in the symbiont shuffling response (Fig. 4B).
+
+![Figure 4.
+Symbiont shuffling as a function of the photochemical advantage of clade D and heat stress duration.
+Panel **(A)** shows the symbiont shuffling response (as in Fig.
+3B) as a function of heat stress duration (days) and the photochemical advantage of clade D under heat stress (A^D^~Heat~), with lines corresponding to the A^D^~Heat~ values for the three coral species studied (indicated by colored lines), and other (unobserved) values (gray lines).
+Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (psuedo-R^2^=0.93) by the time-integrated photochemical advantage of clade D under heat stress.](../figures/Fig4.png)
+
 # Discussion
+
+These experiments revealed significant differences in the photochemical efficiencies of co-occurring symbiont clades in mixed symbiont communities.
+At ambient temperatures, clade D symbionts had significantly lower *F~v~/F~m~* than the other symbiont clade within each coral host.
+Since ambient temperatures are not expected to cause photodamage, the reduced photochemical efficiency of clade D may be due to reduced nutrient sufficiency [@Parkhill:2001eu], greater pigment antenna size or association with PSI [@Suggett:2009p4201], greater excitation 'spillover' from PSII to PSI [@Slavov:2016kh], or other mechanisms of non-photochemical quenching [@Osmond:1994ua].
+Indeed, inherent differences in photobiological traits produce taxonomic variation in *F~v~/F~m~* among broad microalgal classes [@Suggett:2009p4201] as well as different types of *Symbiodinium* [@Suggett:2008p960, @Suggett:2015dv].
+Regardless of the underlying mechanisms, the reduced photochemical efficiency of clade D relative to other *Symbiodinium* *in hospite* suggests clade D has a reduced potential for photochemistry.
+This disadvantage may explain why D-dominated corals tend to grow more slowly [@Little:2004p140], and why clade D may be displaced by other symbionts in the absence of stress [@Thornhill:2006p3638].
+
+However, under heat stress, the relative performance of symbionts was reversed, with clade D maintaining a higher photochemical efficiency than each co-occurring symbiont.
+These reductions in *F~v~/F~m~* can be attributed to accumulation of damage impacting photosystem II, which is known to vary among *Symbiodinium* types [@Warner:1996p8115].
+Indeed, clade D *Symbiodinium* have been shown to maintain higher *F~v~/F~m~* under heat stress [@Rowan:2004p3383, @Silverstein:2015er], indicative of greater thermal tolerance.
+However, the magnitude of clade D's advantage under heat stress differed among the three coral species, being greatest in *M.
+cavernosa*, intermediate in *S.
+siderea*, and lowest in *O.
+faveolata*.
+
+The differential performance of different symbionts in different coral hosts may be explained by variation in symbiont-specific photobiological traits, and/or host-specific traits that modulate symbiont performance.
+For example, coral tissue architecture may influence internal light and nutrient microenvironments [@Wangpraseurt:2015hh], and host regulation of nitrogen and/or carbon dioxide supply to symbionts may influence their performance [@Tansik:2017ct].
+In both *O.
+faveolata* and *S.
+siderea*, clade D symbionts showed the same *F~v~/F~m~*, indicating that the different relative advantages of clade D were in fact due to differential performance of the co-occurring symbionts.
+Clade D in *M.
+cavernosa* had slightly lower *F~v~/F~m~* under ambient conditions (and higher under heat stress) compared to clade D in the other two corals.
+However, since this was repeated bleaching in a separate experiment for *M.
+cavernosa*, we are unable to attribute these differences to unique host traits, the influence of prior bleaching and recovery, or random experimental effects.
+
+Nevertheless, we could test whether the relative photophysiological performance of the two symbionts within each coral host was predictive of the magnitude of symbiont shuffling within that host following heat stress.
+Indeed, 93% of the variability in symbiont shuffling was predicted by *A^D^~Heat~* and the duration of heat stress applied.
+Specifically, greater increases in clade D occurred when *A^D^~Heat~* was higher, and when heat stress was longer.
+Symbiont assemblages in which D had less of an advantage under stress (e.g.
+*O.
+faveolata*) required more stress to induce shuffling to D, while assemblages in which clade D had a large advantage required less stress (e.g.
+*M.
+cavernosa*).
+This finding suggests that both the relative sensitivity to stress and the duration of stress determine symbiont shuffling, similar to the way that both concentration and duration of exposure influence a pharmacological response [@Miller:2000df].
+
+This statistical result reveals predictors but not necessarily the mechanisms of symbiont shuffling.
+However, we speculate that if clade D performs better under at 32°C, it may be expelled at lower rates than co-occurring symbionts [@Silverstein:2017ep], or may even continue to proliferate.
+Furthermore, after heat stress is alleviated (but corals have bleached), amplified internal light environments characteristic of bleached corals [@Wangpraseurt:2012cu] may sustain the photochemical advantage of clade D, allowing it to proliferate at higher rates as the symbiont population regrows.
+Only once symbionts reach high densities may the internal microenvironment (e.g., darker, more nutrient-scarce) again put clade D at a disadvantage, explaining why D may be gradually replaced over time in the absence of stress [@Thornhill:2006p3638].
+Importantly, the rate of change in symbiont community composition in the scenario outlined here (i.e.
+the magnitude of shuffling) would be proportional to the relative difference in the advantage of one symbiont over another, consistent with our observations across coral species.
+
+Whether these same rules may apply in other *Symbiodinium* communities, particularly those without clade D, but which demonstrate shuffling between other types (e.g.
+*Porites* with clades A and C [@Grottoli:2014jo]), remains to be investigated.
+Other populations of *O.
+faveolata* may shuffle variously among clades A, B, C, and D (Jones et al., Serrano et al., unpublished).
+On the other hand, symbiont shuffling appears to be very rare in *Pocillopora* in the eastern Pacific [@McGinley:2012jn] and *Montipora capitata* in Hawaii [@Cunning:2016jb], despite the fact that these corals host members of *Symbiodinium* clades C and D that differ greatly in their sensitivity to stress.
+While the links identified here may not explain symbiont community dynamics in all cases, this study expands our understanding of symbiont shuffling in a variety of important Caribbean reef-building species.
+A better understanding of this phenomenon will allow us to refine predictive models of coral responses to environmental change [@Logan:2013hp], as well as methods to manipulate symbiont communities in corals, which may be an increasingly important tool in coral reef conservation applications [@vanOppen:2015es].
 

### 2017-07-14 manuscript/manuscript.Rmd
AB edits

--- 2017-06-16 manuscript/manuscript.Rmd
@@
 ---
-title: Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
-author: "Ross Cunning"
-date: "6/7/2017"
-output: html_document
+title: Symbiont shuffling linked to differential photochemical dynamics of *Symbiodinium*
+  in three Caribbean reef corals
+author:
+- name: Ross Cunning
+  affiliation: MBE
+  footnote: Corresponding Author
+  email: ross.cunning@gmail.com
+- name: Rachel N. Silverstein
+  affiliation: MBE,MWK
+- name: Andrew C. Baker
+  affiliation: MBE
+
+bibliography: library.bib
+output:
+  rticles::elsevier_article:
+    fig_caption: yes
+  word_document: null
+csl: coral-reefs.csl
+address:
+- address: Department of Marine Biology and Ecology, Rosenstiel School of Marine and
+    Atmospheric Science, University of Miami, 4600 Rickenbacker Causeway, Miami, FL
+    33149, USA
+  code: MBE
+- address: Miami Waterkeeper, 12568 N Kendall Drive, Miami, FL 33185, USA
+  code: MWK
 ---
+```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
+knitr::opts_knit$set(root.dir = "..")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(lmerTest)
+library(MASS)
+```
 
-Title ideas:
+```{r load_analysis}
+load("analysis/analysis.RData")
 
-Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
+ncolsperspp <- with(bleach3c, aggregate(Colony ~ Species, 
+                                        FUN=function(x) length(unique(x))))
+nOfavcol <- ncolsperspp[1, 2]
+nSsidcol <- ncolsperspp[2, 2]
+nMcavcol <- ncolsperspp[3, 2]
+
+ncorepercol <- with(bleach3c, table(Species, Colony))
+ncorepercol[ncorepercol==0] <- NA
+rangeOfav <- range(ncorepercol[1, ], na.rm=T)
+rangeSsid <- range(ncorepercol[2, ], na.rm=T)
+rangeMcav <- range(ncorepercol[3, ], na.rm=T)
+
+ncoreperspp <- with(bleach3c, table(Species))
+nOfavcore <- ncoreperspp["Of"]
+nSsidcore <- ncoreperspp["Ss"]
+nMcavcore <- ncoreperspp["Mc"]
+
+```
+
+# Abstract
+  Dynamic symbioses with functionally diverse dinoflagellate algae in the genus *Symbiodinium* may allow some reef corals to alter their phenotypes through 'symbiont shuffling', or changes in symbiont community composition.
+In particular, corals have been shown to become more bleaching-resistant by increasing the relative abundance of thermally tolerant *Symbiodinium* in clade D after bleaching.
+Despite the immediate relevance of this phenomenon to corals living in warming oceans -- and to interventions aimed at boosting coral resilience -- the mechanisms governing how, why, and when symbiont shuffling occurs are still poorly understood.
+Here, we performed controlled thermal bleaching and recovery experiments on three species of Caribbean reef-building corals, and show that the degree of symbiont shuffling is determined by (1) the difference in photochemical efficiency (*F~v~/F~m~*) of co-occurring symbionts under stress, and (2) the duration of stress exposure.
+The photochemical advantage of D1a (*S.
+trenchii*) under stress was greatest in *Montastraea cavernosa*, intermediate in *Siderastrea siderea*, and lowest in *Orbicella faveolata*, and correlated positively with the magnitude of increase in D1a.
+In holobionts where D1a had less of an advantage (i.e., a smaller difference in *F~v~/F~m~*), longer periods of heat stress were required to elicit commensurate increases in D1a.
+These findings help elucidate general rules governing the process of symbiont shuffling in reef corals, which will help refine predictions of coral responses to environmental change, and inform interventions to manipulate symbiont communities to enhance coral resilience.
+
+# Introduction
+
+Coral reef ecosystems are declining globally at an alarming rate, owing primarily to mass coral bleaching caused by anthropogenic increases in sea surface temperature [@Hughes:2017bs].
+Heat-induced bleaching, or the loss of corals' nutritional symbionts (*Symbiodinium* spp.; @Jokiel:1977p7353), often leads to coral mortality, which precipitates reef ecosystem degradation [@Baker:2008p3421].
+
+However, depending on the severity of bleaching, some corals may recover their *Symbiodinium* communities within weeks to months [@Jokiel:1977p7353; @Jones:1997p3499; @Cunning:2016jb], and ultimately return to a healthy state.
+Moreover, the loss and subsequent re-establishment of the symbiotic algal community may provide an opportunity for change in its composition, i.e., the relative abundance of different *Symbiodinium* types [@Buddemeier:1993p3; @Baker:2001p42].
+When such 'symbiont shuffling' [@Baker:2003p200] occurs, the new configuration of the *Symbiodinium* assemblage may be better suited to tolerate future bleaching stressors [@Berkelmans:2006p191; @Silverstein:2015er], indicating that bleaching may have adaptive value in allowing corals to rapidly acclimatize to changing conditions.
+Moreover, this phenomenon could be leveraged as a way to increase the resilience of reef corals through directed manipulation of *Symbiodinium* communities [@Baker:2008p3421].
+
+*Symbiodinium* is a highly diverse genus comprised of nine clades [@Pochon:2014bk], which are further subdivided into numerous types [with species descriptions underway; @LaJeunesse:2014io; @Wham:2017dy] with varying functional biology [@Suggett:2015dv].
+*S.
+trenchii*, a member of *Symbiodinium* clade D, is a heat-tolerant symbiont that confers a bleaching threshold ~1-2°C higher than other *Symbiodinium* types [@Silverstein:2017ep].
+Increases in the abundance of clade D symbionts have been observed during and after bleaching in several coral species [@Jones:2008p113; @LaJeunesse:2009p7364], and increases in clade A have been observed in others [@Grottoli:2014jo].
+Moreover, the vast majority of coral species surveyed have been observed in association with *Symbiodinium* from multiple clades [@Silverstein:2012ub], suggesting the potential for symbiont shuffling is widespread among corals.
+
+Nevertheless, the mechanisms governing the process of symbiont shuffling, and why it readily occurs in some corals but not others, are still poorly understood.
+In *Orbicella faveolata*, symbiont shuffling was shown to depend on the severity of bleaching (i.e., the magnitude of disturbance to the existing community), as well as the environmental conditions during recovery [@Cunning:2015ja], suggesting that the relative performance of different symbionts influences the trajectory of symbiont community re-establishment.
+However, general rules that link symbiont performance to variability in symbiont shuffling among coral species with different *Symbiodinium* types have never been identified.
+Ultimately, a deeper understanding of symbiont community dynamics in corals is a critical foundation for 1) predicting corals responses to environmental change [@Logan:2013hp], and 2) manipulating symbiont communities as a way to increase resilience in coral reef restoration [@vanOppen:2015es].
+
+Here, we investigated symbiont shuffling through controlled bleaching experiments with three Caribbean coral species hosting mixed *Symbiodinium* assemblages: *O.
+faveolata* (with clades B and D), *S.
+siderea* (with clades C and D), and *M.
+cavernosa* (with clades C and D).
+We used chlorophyll fluorescence in combination with clade-specific quantitative PCR assays to explore links between the photochemical efficiencies of co-occurring symbionts and the magnitude of symbiont shuffling observed after bleaching.
+
+# Materials and Methods
+
+## Experimental manipulations
+
+Two separate bleaching and recovery experiments were conducted.
+The first experiment included both *O.
+faveolata* (`r nOfavcol` colonies, `r min(rangeOfav)`-`r max(rangeOfav)` cores colony^-1^, n=`r nOfavcore` cores) and *S.
+siderea* (`r nSsidcol` colonies, `r min(rangeSsid)`-`r max(rangeSsid)` cores colony^-1^, n=`r nSsidcore` cores), with bleaching treaments of exposure to 32°C for 7, 10, and 14 days [@Cunning:2015ja], and recovery for 3 months at 24 or 29°C.
+The second experiment was conducted on *M.
+cavernosa* (`r nMcavcol` colonies, `r min(rangeMcav)`-`r max(rangeMcav)` cores colony^-1^, n=`r nMcavcore` cores), with a bleaching treatment of 32°C for 10 days [@Silverstein:2015er] and recovery for 3 months at 24 or 29°C.
+Prior to experimentation, the *O.
+faveolata* cores contained varying mixtures of *Symbiodinium* in clades B and D [@Cunning:2015ja], while *S.
+siderea* colonies contained mixtures of clades C and D [@Cunning:2013tl].
+*M.
+cavernosa* contained only clade C *Symbiodinium* upon collection, but an initial bleaching and recovery experiment involving some of these cores produced a range of mixtures of clades C and D [@Silverstein:2015er].
+Both experiments were conducted in the same indoor, semi-recirculating coral culture facility at the University of Miami.
+For additional details on coral collection and experimental set-up, see @Silverstein:2015er and @Cunning:2015ja.
+
+## Sampling and data collection
+
+In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of heat stress using an Imaging PAM (Walz, Effeltrich, Germany).
+At the same time points, small tissue biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR.
+Tissue samples were collected again after 3 months of recovery from bleaching to analyze changes in *Symbiodinium* community structure (i.e., symbiont shuffling).
+For additional details on sampling and data collection, see @Silverstein:2015er and @Cunning:2015ja.
+
+## Data analysis
+
+Photochemical efficiency (*F~v~/F~m~*) was analyzed by a linear mixed model (*lmer*; [@Bates:2015wh]) with fixed effects of time (pre- or post-heat-stress), proportion D1a (measured at the same time point), and coral species, and with random intercepts for each colony.
+Backward elimination of non-significant effects from a fully crossed model was performed using partial F-tests in *lmerTest* [@lmerTestTestsinL:j0go4ZGL].
+By relating the photochemical efficiency of *Symbiodinium* communities to their taxonomic composition [*sensu*  @Suggett:2009p4201], we can tease apart the relative photochemical performance of the two different symbionts in each coral.
+Specifically, regression coefficients for the proportion D1a were interpreted as the photochemical advantage of D1a for each species under ambient (pre-heat-stress; *A^D^~Amb~*) and heated (post-heat-stress; *A^D^~Heat~*) conditions, and tested for statistical differences by pairwise comparisons using *lsmeans* [@Lenth:2015vl].
+Stress duration (i.e., 7, 10, or 14 days) was not included as a predictor since it had no impact on the photochemical advantage of D1a in the first experiment (i.e., no significant interaction between stress duration and proportion D1a).
+
+Symbiont shuffling was analyzed by modeling the final proportion of D1a (after 3 months of recovery) as a function of the initial proportion D1a (before bleaching), using a quasi-binomial generalized linear model (GLM) with species, stress duration, and recovery temperature as co-predictors.
+Fitted values for the final proportion of D1a for each species at each stress duration (averaged across recovery temperatures, and predicted for *M.
+cavernosa* at 7- and 14-day durations using the *effects* package [@Fox:2003uu]), were plotted against initial proportion of D1a with 90% confidence intervals.
+From these fitted values, an integrated metric of symbiont shuffling was generated across the full range of initial D1a proportions by finding the area between the fitted curve for each group and the diagonal line of identity, which indicates no change in symbiont proportions.
+This metric was scaled to a range of -1 (e.g., complete loss of D1a) to +1 (e.g., complete dominance by D1a), with zero indicating no symbiont shuffling.
+
+Finally, we used a quasi-binomial GLM to model the integrated symbiont shuffling response as a function of the photochemical advantage of D1a under heat stress (A^D^~Heat~) for each species, and days of exposure to heat stress.
+All data analysis was conducted using R v3.3.1 [@RALanguageandEn:2014wf].
+Data and R code to reproduce all analyses and figures presented here are available at: github.com/jrcunning/symshuff-3c.
+
+# Results
+
+## Photochemistry of symbiont assemblages
+
+Measured values of *F~v~/F~m~* were significantly affected by the proportion of D1a in symbiont assemblages (Fig.
+1), but this effect varied among coral species and between ambient and heat-stress conditions (three-way interaction, *F*=`r round(lmerTest::anova(bleachmod)$F[7], 3)`, *p*=`r round(lmerTest::anova(bleachmod)$P[7], 3)`).
+In all three coral species, *F~v~/F~m~* was negatively related to the proportion of D1a under ambient conditions, but positively related to the proportion of D1a under heat stress.
+The slope of each line in Fig.
+1 (i.e., the difference in *F~v~/F~m~* when proportion D=1 and when proportion D=0) can be interpreted as the photochemical advantage of D1a compared to the other symbiont (Fig.
+2).
+The advantage of D under ambient conditions (*A^D^~Amb~*) was significantly more negative (i.e., a greater disadvantage) in *O.
+faveolata* (`r round(AD[4,3], 3)`) compared to *M.
+cavernosa* (`r round(AD[6,3], 3)`, *p*=`r round(lst.pw$p.value[5], 3)`), and intermediate (`r round(AD[5,3], 3)`, but not significantly different, *p*>`r round(min(lst.pw$p.value[c(4,6)]), digits=1)`) in *S.
+siderea*.
+Under heat stress, the advantage of D1a (*A^D^~Heat~*) was greatest in *M.
+cavernosa* (`r round(AD[3,3], 3)`), intermediate in *S.
+siderea* (`r round(AD[2,3], 3)`), and lowest in *O.
+faveolata* (`r round(AD[1,3], 3)`, all pairwise comparisons, *p*<`r plyr::round_any(max(lst.pw$p.value[c(1,2,3)]), 0.001, ceiling)`).
+
+![Figure 1.
+Photochemical efficiency as a function of the proportion of D1a *Symbiodinium* in **A)** *O.
+faveolata*, **B)** *S.
+siderea*, and **C)** *M.
+cavernosa*.
+Solid lines represent ambient conditions, and dotted lines represent the relationship under heat stress (32$^\circ$C).
+Shaded regions represent 90% confidence intervals around fitted values.](../figures/Fig1.png)
+
+![Figure 2.
+Photochemical advantage of D1a symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars).
+Values indicate the difference in Fv/Fm of corals with all and no D1a symbionts (i.e., slopes of the lines in Fig.
+1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)
+
+## Symbiont shuffling
+
+The proportion of D1a after recovery from bleaching (Fig.
+3A) depended on the initial proportion of D1a (*F*=`r round(dropterm(propDmod, test="F")$F[2], 3)`, *p*`r format.pval(dropterm(propDmod, test="F")$P[2], eps=0.0001)`) and an interaction among coral species, heat stress duration, and recovery temperature (*F*=`r round(dropterm(propDmod, test="F")$F[3], 3)`, *p*=`r round(dropterm(propDmod, test="F")$P[3], 3)`).
+Fitted values were averaged across recovery temperatures in order to focus on differences among species and heat stress durations.
+Across the full range of initial D1a proportions, the integrated symbiont shuffling response, $S$ (Fig.
+3B), showed a loss of D1a in *O.
+faveolata* after recovery from 7 and 10 days of heat stress ($S$=`r round(auc[4,3], 3)` and `r round(auc[7,3], 3)`), but a gain of D1a after recovery from 14 days of heat stress ($S$=`r round(auc[1,3], 3)`).
+In contrast, *S.
+siderea* underwent greater shuffling toward D1a after 7 ($S$=`r round(auc[5,3], 3)`), 10 ($S$=`r round(auc[8,3], 3)`), and 14 ($S$=`r round(auc[2,3], 3)`) days of heat stress.
+*M.
+cavernosa* shuffled almost completely to D1a assemblages following 10 days of stress ($S$=`r round(auc[9,3], 3)`), and thus the statistical model also predicted near complete shuffling to D after 7 ($S$=`r round(auc[6,3], 3)`) and 14 ($S$=`r round(auc[3,3], 3)`) days of heat stress.
+
+![Figure 3.
+Symbiont shuffling in three coral species.
+Panel **(A)** shows the proportion of D1a *Symbiodinium* before bleaching and after recovery, with the diagonal line of identity representing no change in symbiont proportions.
+Fitted values for each species (indicated by color) and heat stress duration (indicated by line type) are plotted with shaded areas representing 95% confidence intervals.
+Panel **(B)** shows the integrated symbiont shuffling response (from before bleaching to after recovery) for each species and bleaching treatment, calculated as the area above or below the diagonal line of identity, normalized to a range of -1 (i.e., a complete loss of D1a symbionts) to +1 (i.e., complete dominance by D1a symbionts).](../figures/Fig3.png)
+
+## Relationship between photochemistry and symbiont shuffling
+
+The magnitude of symbiont shuffling showed a significant statistical relationship (Fig.
+4A) with both the photochemical advantage of D1a under heat stress (*A^D^~Heat~*, *F*=`r round(anova(glmod, test="F")$F[2], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[2], eps=0.0001)`) and the duration of heat stress (*F*=`r round(anova(glmod, test="F")$F[3], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[3], eps=0.0001)`).
+Specifically, increases in the amount of D1a were greater for higher values of *A^D^~Heat~* and longer durations of heat stress.
+The model predicts the duration of heat stress required to cause symbiont shuffling toward D1a for different values of *A^D^~Heat~* (Fig.
+4A), and shows that 7 days of heat stress results in shuffling toward D1a for *A^D^~Heat~*>~0.06.
+Lower values of *A^D^~Heat~* cause a decrease in the proportion of D1a after recovery, even following 14 days of heat stress for *A^D^~Heat~*<~0.01.
+
+Across these three coral species, the interaction of *A^D^~Heat~* and the duration of heat stress explains `r round(pseudoR2*100, 1)`% of the variability in the symbiont shuffling response (Fig. 4B).
+
+![Figure 4.
+Symbiont shuffling as a function of the photochemical advantage of D1a and heat stress duration.
+Panel **(A)** shows the symbiont shuffling response (as in Fig.
+3B) as a function of heat stress duration (days) and the photochemical advantage of D1a under heat stress (A^D^~Heat~), with lines corresponding to the A^D^~Heat~ values for the three coral species studied (indicated by colored lines), and other (unobserved) values (gray lines).
+Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (pseudo-R^2^=0.93) by the time-integrated photochemical advantage of D1a under heat stress.](../figures/Fig4.png)
+
+# Discussion
+
+These experiments revealed significant differences in the photochemical efficiencies of co-occurring symbiont clades in mixed symbiont communities.
+At ambient temperatures, D1a symbionts had significantly lower *F~v~/F~m~* than the other symbiont clade within each coral host.
+Since ambient temperatures are not expected to cause photodamage, the reduced photochemical efficiency of D1a may be due to reduced nutrient sufficiency [@Parkhill:2001eu], greater pigment antenna size or association with PSI [@Suggett:2009p4201], greater excitation 'spillover' from PSII to PSI [@Slavov:2016kh], or other mechanisms of non-photochemical quenching [@Osmond:1994ua].
+Indeed, inherent differences in photobiological traits produce taxonomic variation in *F~v~/F~m~* among broad microalgal classes [@Suggett:2009p4201] as well as different types of *Symbiodinium* [@Suggett:2008p960, @Suggett:2015dv].
+Regardless of the underlying mechanisms, the reduced photochemical efficiency of D1a relative to other *Symbiodinium* *in hospite* suggests D1a has a reduced potential for PSII photochemistry.
+This disadvantage may explain why D-dominated corals tend to grow more slowly [@Little:2004p140], and why D1a may be displaced by other symbionts in the absence of stress [@Thornhill:2006p3638; @LaJeunesse:2009p7364].
+
+However, under heat stress, the relative performance of symbionts was reversed, with D1a maintaining a higher photochemical efficiency than each co-occurring symbiont.
+These reductions in *F~v~/F~m~* can be attributed to accumulation of damage impacting photosystem II, which is known to vary among *Symbiodinium* types [@Warner:1996p8115].
+Indeed, clade D *Symbiodinium* have been shown to maintain higher *F~v~/F~m~* under heat stress [@Rowan:2004p3383, @Silverstein:2015er], indicative of greater thermal tolerance.
+However, the magnitude of D1a's advantage under heat stress differed among the three coral species, being greatest in *M.
+cavernosa*, intermediate in *S.
+siderea*, and lowest in *O.
+faveolata*.
+
+The differential performance of different symbionts in different coral hosts may be explained by variation in symbiont-specific photobiological traits, and/or host-specific traits that modulate symbiont performance.
+For example, coral tissue architecture may influence internal light and nutrient microenvironments [@Wangpraseurt:2015hh], and host regulation of nitrogen and/or carbon dioxide supply to symbionts may influence their performance [@Tansik:2017ct].
+In both *O.
+faveolata* and *S.
+siderea*, D1a symbionts showed the same *F~v~/F~m~*, indicating that the different relative advantages of D1a were in fact due to differential performance of the co-occurring symbionts.
+D1a in *M.
+cavernosa* had slightly lower *F~v~/F~m~* under ambient conditions (and higher under heat stress) compared to D1a in the other two corals.
+However, since this followed repeated bleaching in a separate experiment for *M.
+cavernosa*, we are unable to attribute these differences to unique host traits, the influence of prior bleaching and recovery, or random experimental effects.
+
+Nevertheless, we could test whether the relative photophysiological performance of the two symbionts within each coral host was predictive of the magnitude of symbiont shuffling within that host following heat stress.
+Indeed, 93% of the variability in symbiont shuffling was predicted by *A^D^~Heat~* and the duration of heat stress applied.
+Specifically, greater increases in D1a occurred when *A^D^~Heat~* was higher, and when heat stress was longer.
+Symbiont assemblages in which D had less of an advantage under stress (e.g.
+*O.
+faveolata*) required more stress to induce shuffling to D1a, while assemblages in which D1a had a large advantage (e.g.
+*M.
+cavernosa*) required less stress.
+This finding suggests that both the relative sensitivity to stress and the duration of stress determine symbiont shuffling, similar to the way that both concentration and duration of exposure influence a pharmacological response [@Miller:2000df].
+
+This statistical result reveals predictors of symbiont shuffling but not necessarily the mechanisms underlying them.
+However, we speculate that if D1a performs better under at 32°C, it may be expelled at lower rates than co-occurring symbionts [@Silverstein:2017ep], or may even continue to proliferate.
+Furthermore, after heat stress is alleviated (but corals have bleached), amplified internal light environments characteristic of bleached corals [@Wangpraseurt:2012cu] may sustain the photochemical advantage of D1a, allowing it to proliferate at higher rates as the symbiont population regrows.
+Only once symbionts reach high densities may the internal microenvironment (e.g., darker, more nutrient-scarce conditions) again put D1a at a disadvantage, explaining why D may be gradually replaced over time in the absence of stress [@Thornhill:2006p3638].
+Importantly, the rate of compositional change in the scenario outlined here would be proportional to the relative difference in the advantage of one symbiont over another, consistent with our observations across coral species.
+
+Whether these same rules may apply in other *Symbiodinium* communities, particularly those without clade D, but which demonstrate shuffling between other types [e.g., *Porites* with clades A and C; @Grottoli:2014jo], remains to be investigated.
+Other populations of *O.
+faveolata* may shuffle variously among clades A, B, C, and D [@Baker:1999p147; Jones et al., Serrano et al., unpublished).
+In contrast, symbiont shuffling appears to be rare in other coral taxa, such as *Pocillopora* in the Mexican Pacific [@McGinley:2012jn] and *Montipora capitata* in Hawaii [@Cunning:2016jb], despite the fact that these corals host members of *Symbiodinium* clades C and D that differ greatly in their sensitivity to stress.
+While the links identified here may not explain symbiont community dynamics in all cases, this study expands our understanding of symbiont shuffling in a variety of important Caribbean reef-building species.
+A better understanding of this phenomenon will allow us to refine predictive models of coral responses to environmental change [@Logan:2013hp], as well as methods to manipulate symbiont communities in corals, which may be an increasingly important tool in coral reef conservation [@vanOppen:2015es; @Aswani:2015cq].
+
+# Acknowledgements
+RC and RNS were supported by NSF Graduate Research Fellowships and University of Miami Graduate Fellowships, with additional funding from RSMAS student awards, the Captain Harry D.
+Vernon Memorial Scholarship, and the Garden Club of America.
+Additional support was provided by NSF (OCE-0547169 and OCE-1358699), the Lenfest Ocean Program, a Pew Fellowship in Marine Conservation, and a University of Miami Provost Research Award in Natural Science and Engineering to ACB.
+We thank Z Schwartz, K Ondrasik, K Dziedzic, N Formel, K Montenero, P Jones, R Winter, R Okazaki, L Gordon, and N.
+Guy for laboratory assistance.
+
+# References

### 2017-07-18 manuscript/manuscript.Rmd
RS edits

@@
 author:
-- name: Ross Cunning
-  affiliation: MBE
+- affiliation: MBE
+  email: ross.cunning@gmail.com
   footnote: Corresponding Author
-  email: ross.cunning@gmail.com
-- name: Rachel N. Silverstein
-  affiliation: MBE,MWK
-- name: Andrew C. Baker
-  affiliation: MBE
-
-bibliography: library.bib
-output:
-  rticles::elsevier_article:
-    fig_caption: yes
-  word_document: null
-csl: coral-reefs.csl
+  name: Ross Cunning
+- affiliation: MBE,MWK
+  name: Rachel N. Silverstein
+- affiliation: MBE
+  name: Andrew C. Baker
 address:
@@
   code: MWK
+output:
+  word_document: null
+  rticles::elsevier_article:
+    fig_caption: yes
+csl: coral-reefs.csl
+bibliography: library.bib
+abstract: |
+  Dynamic symbioses with functionally diverse dinoflagellate algae in the genus *Symbiodinium* may allow some reef corals to alter their phenotypes through 'symbiont shuffling', or changes in symbiont community composition.
+In particular, corals have been shown to become more bleaching-resistant by increasing the relative abundance of thermally tolerant *Symbiodinium* in clade D after bleaching.
+Despite the immediate relevance of this phenomenon to corals living in warming oceans -- and to interventions aimed at boosting coral resilience -- the mechanisms governing how, why, and when symbiont shuffling occurs are still poorly understood.
+Here, we performed controlled thermal bleaching and recovery experiments on three species of Caribbean reef-building corals, and show that the degree of symbiont shuffling is determined by (1) the difference in photochemical efficiency (*F~v~/F~m~*) of co-occurring symbionts under stress, and (2) the duration of stress exposure.
+The photochemical advantage of D1a (*S.
+trenchii*) under stress was greatest in *Montastraea cavernosa*, intermediate in *Siderastrea siderea*, and lowest in *Orbicella faveolata*, and correlated positively with the magnitude of shuffling toward D1a.
+In holobionts where D1a had less of an advantage under stress (i.e., only slightly higher *F~v~/F~m~*), longer periods of stress were required to elicit commensurate increases in D1a.
+These findings help elucidate general rules governing the process of symbiont shuffling in reef corals, which will help refine predictions of coral responses to environmental change and inform interventions to manipulate symbiont communities to enhance coral resilience.
 ---
+
 ```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-# Abstract
-  Dynamic symbioses with functionally diverse dinoflagellate algae in the genus *Symbiodinium* may allow some reef corals to alter their phenotypes through 'symbiont shuffling', or changes in symbiont community composition.
-In particular, corals have been shown to become more bleaching-resistant by increasing the relative abundance of thermally tolerant *Symbiodinium* in clade D after bleaching.
-Despite the immediate relevance of this phenomenon to corals living in warming oceans -- and to interventions aimed at boosting coral resilience -- the mechanisms governing how, why, and when symbiont shuffling occurs are still poorly understood.
-Here, we performed controlled thermal bleaching and recovery experiments on three species of Caribbean reef-building corals, and show that the degree of symbiont shuffling is determined by (1) the difference in photochemical efficiency (*F~v~/F~m~*) of co-occurring symbionts under stress, and (2) the duration of stress exposure.
-The photochemical advantage of D1a (*S.
-trenchii*) under stress was greatest in *Montastraea cavernosa*, intermediate in *Siderastrea siderea*, and lowest in *Orbicella faveolata*, and correlated positively with the magnitude of increase in D1a.
-In holobionts where D1a had less of an advantage (i.e., a smaller difference in *F~v~/F~m~*), longer periods of heat stress were required to elicit commensurate increases in D1a.
-These findings help elucidate general rules governing the process of symbiont shuffling in reef corals, which will help refine predictions of coral responses to environmental change, and inform interventions to manipulate symbiont communities to enhance coral resilience.
-
 # Introduction
 
-Coral reef ecosystems are declining globally at an alarming rate, owing primarily to mass coral bleaching caused by anthropogenic increases in sea surface temperature [@Hughes:2017bs].
+Coral reef ecosystems are declining globally at an alarming rate, primarily due to mass coral bleaching caused by anthropogenic increases in sea surface temperature [@Hughes:2017bs].
 Heat-induced bleaching, or the loss of corals' nutritional symbionts (*Symbiodinium* spp.; @Jokiel:1977p7353), often leads to coral mortality, which precipitates reef ecosystem degradation [@Baker:2008p3421].
@@
 
-In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of heat stress using an Imaging PAM (Walz, Effeltrich, Germany).
+In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of heat stress using an Imaging PAM fluorometer (Walz, Effeltrich, Germany).
 At the same time points, small tissue biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR.
@@
 1), but this effect varied among coral species and between ambient and heat-stress conditions (three-way interaction, *F*=`r round(lmerTest::anova(bleachmod)$F[7], 3)`, *p*=`r round(lmerTest::anova(bleachmod)$P[7], 3)`).
-In all three coral species, *F~v~/F~m~* was negatively related to the proportion of D1a under ambient conditions, but positively related to the proportion of D1a under heat stress.
+In all three coral species, a higher relative abundance of D1a led to lower *F~v~/F~m~* under ambient conditions, but higher *F~v~/F~m~* under heat stress.
 The slope of each line in Fig.
@@
 
-These experiments revealed significant differences in the photochemical efficiencies of co-occurring symbiont clades in mixed symbiont communities.
+These experiments revealed differences in the degree of symbiont shuffling in three coral species, which were strongly related to differences in the photochemical efficiencies of co-occurring symbionts.
 At ambient temperatures, D1a symbionts had significantly lower *F~v~/F~m~* than the other symbiont clade within each coral host.
@@
 Indeed, inherent differences in photobiological traits produce taxonomic variation in *F~v~/F~m~* among broad microalgal classes [@Suggett:2009p4201] as well as different types of *Symbiodinium* [@Suggett:2008p960, @Suggett:2015dv].
-Regardless of the underlying mechanisms, the reduced photochemical efficiency of D1a relative to other *Symbiodinium* *in hospite* suggests D1a has a reduced potential for PSII photochemistry.
+Regardless of the underlying mechanisms, the reduced photochemical efficiency of D1a relative to other *Symbiodinium* *in hospite* suggests D1a has a reduced potential for PSII photochemistry under ambient conditions.
 This disadvantage may explain why D-dominated corals tend to grow more slowly [@Little:2004p140], and why D1a may be displaced by other symbionts in the absence of stress [@Thornhill:2006p3638; @LaJeunesse:2009p7364].
@@
 Furthermore, after heat stress is alleviated (but corals have bleached), amplified internal light environments characteristic of bleached corals [@Wangpraseurt:2012cu] may sustain the photochemical advantage of D1a, allowing it to proliferate at higher rates as the symbiont population regrows.
-Only once symbionts reach high densities may the internal microenvironment (e.g., darker, more nutrient-scarce conditions) again put D1a at a disadvantage, explaining why D may be gradually replaced over time in the absence of stress [@Thornhill:2006p3638].
+Only once symbionts reach high densities may the internal microenvironment (e.g., darker, more nutrient-scarce conditions) again put D1a at a disadvantage, explaining why clade D may be gradually replaced over time in the absence of stress [@Thornhill:2006p3638].
 Importantly, the rate of compositional change in the scenario outlined here would be proportional to the relative difference in the advantage of one symbiont over another, consistent with our observations across coral species.

### 2017-10-06 analysis/analysis.Rmd
revisions

@@
 plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
-     xlab="Prop. clade D", ylab="Fv/Fm")
+     xlab="Proportion D1a", ylab=expression(italic(F[v]/F[m])))
 mtext("A.", adj=-0.2, font=2)
@@
 plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
-     xlab="Prop. clade D", ylab="")
+     xlab="Proportion D1a", ylab="")
 mtext("B.", adj=-0.2, font=2)
@@
 plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
-     xlab="Prop. clade D", ylab="")
+     xlab="Proportion D1a", ylab="")
 mtext("C.", adj=-0.2, font=2)
@@
        length = 0.05, lwd=1, xpd=NA, code=3, angle=90)
-mtext(side=2, "Photochemical advantage of clade D", 
+mtext(side=2, "Photochemical advantage of D1a", 
       line=1.5, cex=0.8)
@@
 ![Figure 2.
-Photochemical advantage of clade D symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars).
-Values indicate the difference in Fv/Fm of corals with all and no clade D symbionts (i.e., slopes of the lines in Fig.
+Photochemical advantage of D1a symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars).
+Values indicate the difference in Fv/Fm of corals with all and no D1a symbionts (i.e., slopes of the lines in Fig.
 1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)
@@
 plot(NA, xlim=c(0,1), ylim=c(0,1), xaxs="i", yaxs="i",
-     xlab="Prop. D before bleaching",
-     ylab="Prop. D after recovery")
+     xlab="Prop. D1a before bleaching",
+     ylab="Prop. D1a after recovery")
 # Plot lines for low bleaching
@@
 text(par("usr")[2]*0.95, c(0.08, -0.1), pos=2, cex=0.7,
-     labels=c("more clade D", "less clade D"))
+     labels=c("more D1a", "less D1a"))
 #text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "B.", font=2, xpd=NA)
@@
 
-# P#d95f02ict shuffling response based on relative Fv/Fm value and BleachLvl
+# Predict shuffling response based on relative Fv/Fm value and BleachLvl
 ## relative Fv/Fm at ambient temperatures ("normal") is perfectly correlated with relative Fv/Fm at stress temperatures ("stress") so only one variable can be used in model.

### 2017-11-01 analysis/analysis.Rmd
Figure adjustments for publication

@@
      xlab="Proportion D1a", ylab=expression(italic(F[v]/F[m])))
-mtext("A.", adj=-0.2, font=2)
+mtext("a", adj=-0.2, font=2)
 with(subset(pred, Species=="Of" & time=="init"), {
@@
      xlab="Proportion D1a", ylab="")
-mtext("B.", adj=-0.2, font=2)
+mtext("b", adj=-0.2, font=2)
 with(subset(pred, Species=="Ss" & time=="init"), {
@@
      xlab="Proportion D1a", ylab="")
-mtext("C.", adj=-0.2, font=2)
+mtext("c", adj=-0.2, font=2)
 with(subset(pred, Species=="Mc" & time=="init"), {
@@
 #text(-0.2, par("usr")[4], "A.", font=2, xpd=NA)
-mtext("A.", adj=-0.2, font=2)
+mtext("a", adj=-0.2, font=2)
 
@@
      xlab="Species", ylab="Symbiont shuffling")
-axis(side=1, at=1:3, labels=c("Of", "Ss", "Mc"))
+axis(side=1, at=1:3, labels=c("Ofav", "Ssid", "Mcav"))
 abline(h=0, lty=1, lwd=0.5)
@@
 #text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "B.", font=2, xpd=NA)
-mtext("B.", adj=-0.2, font=2)
+mtext("b", adj=-0.2, font=2)
 dev.off()
@@
      ylab="Symbiont shuffling")
-mtext("A.", adj=-0.2, font=2)
+mtext("a", adj=-0.2, font=2)
 reslist <- split(resdf, f=resdf$ADH)
@@
 mtext("Symbiont shuffling", side=2, line=1.5)
-mtext("B.", adj=-0.2, font=2)
+mtext("b", adj=-0.2, font=2)
 lines(seq(0,3,0.1), rescale(predicted, from=c(0,1), to=c(-1,1)), lwd=1)

### 2017-11-01 analysis/analysis.Rmd
tidy

@@
 ---
-title: "Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals"
+title: "Symbiont shuffling linked to differential photochemical dynamics of *Symbiodinium* in three Caribbean reef corals"
 output: html_document
@@
 Mc$BleachLvl[Mc$BleachLvl=="Heat"] <- "Med"
-## Set data after second recovery to the "recove#d95f02" factor levels
+## Set data after second recovery to the "recovered" factor levels
 Mc$PropD.recov <- Mc$PropD.recov2
@@
 
+## Plot Figure 2
 ```{r Fig2}
@@
 
+## Plot Figure 3
 ```{r Fig3}
@@
 
+## Plot Figure 4
 ```{r Fig4}
@@
 # Predict shuffling response based on relative Fv/Fm value and BleachLvl
-## relative Fv/Fm at ambient temperatures ("normal") is perfectly correlated with relative Fv/Fm at stress temperatures ("stress") so only one variable can be used in model.
-Or, should the slope of the Fv/Fm line be used, rather than the "relative difference"?
-(makes very little difference whether relative or slope is used.) *slope needs to be used because relative is different for diff BleachLvl levels, but slope is the same!
 
